---
# No title/toc here: the title block and table of contents are typeset by the
# title-block chunk below so the first page can carry the logo and permit.
format:
  pdf:
    toc: false
    toc-depth: 1
    number-sections: true
    geometry:
      - margin=1in
      - headheight=30pt
      - headsep=16pt
    documentclass: article
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{float}
        \usepackage{needspace}
        \usepackage{graphicx}
        \usepackage{fancyhdr}
        \usepackage{lastpage}
        \usepackage{sectsty}
        % Colours from www/epa_logo.png. Text only ever uses epablue (or white
        % on epablue): it prints as dark grey in black and white. epagreen is
        % used for thin decorative rules only, never behind or as text.
        \definecolor{epablue}{HTML}{0D6CB6}
        \definecolor{epagreen}{HTML}{46AA42}
        \definecolor{inkgray}{HTML}{333333}
        % Darker orange for the "Note" / "Manual WQS Applied" labels: xcolor's
        % default orange turns pale grey in black-and-white print.
        \definecolor{orange}{HTML}{B35900}
        \allsectionsfont{\sffamily\color{epablue}}
        \AtBeginDocument{\hypersetup{colorlinks=true, linkcolor=epablue, urlcolor=epablue}}
        % Set by the title-block chunk; used by the running header/footer.
        \providecommand{\rppermit}{}
        \providecommand{\rpgenerated}{}
        \newcommand{\rplogo}[1]{\IfFileExists{epa_logo.png}{\includegraphics[height=#1]{epa_logo.png}}{}}
        \pagestyle{fancy}
        \fancyhf{}
        \fancyhead[L]{\raisebox{-0.35\height}{\rplogo{0.34in}}\hspace{6pt}\textcolor{inkgray}{\sffamily\small U.S. EPA Region 2 \textbar{} NPDES Reasonable Potential Report}}
        \fancyhead[R]{\textcolor{inkgray}{\sffamily\small\bfseries \rppermit}}
        \fancyfoot[L]{\textcolor{inkgray}{\sffamily\footnotesize Generated \rpgenerated}}
        \fancyfoot[R]{\textcolor{inkgray}{\sffamily\footnotesize Page \thepage\ of \pageref*{LastPage}}}
        \renewcommand{\headrulewidth}{1pt}
        \renewcommand{\headrule}{\hbox to\headwidth{\color{epablue}\leaders\hrule height \headrulewidth\hfill}}
        \renewcommand{\footrulewidth}{0.6pt}
        \renewcommand{\footrule}{\hbox to\headwidth{\color{epagreen}\leaders\hrule height \footrulewidth\hfill}\vskip 4pt}
        % First page: the title block carries the logo, so footer only.
        \fancypagestyle{rptitle}{\fancyhf{}\renewcommand{\headrulewidth}{0pt}\renewcommand{\headrule}{}\fancyfoot[R]{\textcolor{inkgray}{\sffamily\footnotesize Page \thepage\ of \pageref*{LastPage}}}}
execute:
  echo: false
params:
  permit_id: ""
  facility_name: ""
  date_start: ""
  date_end: ""
  forms: ""
  dilution_ratio: 1
  confidence_level: 0.95
  target_percentile: 0.95
  hardness_mode: "range"
  hardness_value: NA
  rp_path: ""
  coverage_path: ""
  dmr_path: ""
  ph_dmr_path: ""
  temp_dmr_path: ""
  wqs_info_path: ""
  overrides_path: ""
  quick_run: false
  direct_dmr_path: ""    # observations for all rp_method == "direct" parameters
  quick_run_params: ""   # comma-separated list of auto-excluded parameter descs
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = "center", fig.show = "asis",
  # fig.width matches the 6.5in usable text width (letter, 1in margins) so
  # plots render at the correct aspect ratio and never exceed \linewidth.
  # out.width caps the displayed width; fig.height keeps plots compact.
  fig.width = 6.5, fig.height = 3.6, out.width = "\\linewidth"
)

library(readr)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggtext)
library(lubridate)
library(ggforce)
library(ggnewscale)
library(tidyr)

# Inline copy of format_water_class() — keeps the report self-contained when
# rendered from tmp_dir (functions.R is not copied there). If the helper in
# functions.R is updated, mirror the change here.
format_water_class <- function(class, sub_class = NA, sep = " — ") {
  has_sub <- !is.na(sub_class) & nzchar(as.character(sub_class))
  ifelse(has_sub,
         paste0(as.character(class), sep, as.character(sub_class)),
         as.character(class))
}

# Inline copy of format_outfall_label() — same self-containment reason as
# above. Produces "Internal Outfall NNN" or "External Outfall NNN" so the
# type is always visible in the report. Mirror functions.R if this changes.
format_outfall_label <- function(perm_feature_nmbr, perm_feature_type_code = NA_character_) {
  is_internal <- !is.na(perm_feature_type_code) & perm_feature_type_code == "INO"
  ifelse(
    is_internal,
    paste0("Internal Outfall ", perm_feature_nmbr),
    paste0("External Outfall ", perm_feature_nmbr)
  )
}

# LaTeX version: wraps Internal Outfall labels in \textcolor{red}{...} so
# they appear in red in PDF section headings and plot titles.
# External Outfall labels are returned as-is (no colour command needed).
format_outfall_label_latex <- function(perm_feature_nmbr, perm_feature_type_code = NA_character_) {
  lbl <- format_outfall_label(perm_feature_nmbr, perm_feature_type_code)
  is_internal <- !is.na(perm_feature_type_code) & perm_feature_type_code == "INO"
  ifelse(
    is_internal,
    paste0("\\textcolor{red}{", lbl, "}"),
    lbl
  )
}

rp_df       <- readr::read_csv(params$rp_path, show_col_types = FALSE) %>%
  dplyr::mutate(CRITERION_ID = as.character(CRITERION_ID))

# Backfill the legacy Method column if the upstream app no longer supplies it.
# The pH section keys its Min/Max criterion lines on Method; rp_operator is
# the authoritative replacement ("<" = floor = Min bound, ">" = ceiling =
# Max bound). Deriving it here keeps the report renderable against either
# vintage of rp_concentration.csv.
if (!"Method" %in% names(rp_df)) {
  rp_df <- rp_df %>%
    dplyr::mutate(
      Method = if ("rp_operator" %in% names(.)) {
        dplyr::case_when(
          is.na(rp_operator) ~ "Concentration",
          rp_operator == "<" ~ "Min",
          rp_operator == ">" ~ "Max",
          TRUE ~ "Concentration"
        )
      } else {
        "Concentration"
      }
    )
}
coverage_df <- if (nzchar(params$coverage_path)) readr::read_csv(params$coverage_path, show_col_types = FALSE) else NULL
dmr_df      <- if (nzchar(params$dmr_path))      readr::read_csv(params$dmr_path,      show_col_types = FALSE) else NULL
# ── crosswalk_v2 helpers (inline; mirror functions.R if changed) ────────────
# Binary criteria store CRITERION_VALUE = 0 so the >-comparison works. They
# are operationalised narrative standards (PRWQS 1303.1.H oil & grease,
# Class SG surfactants), and print as 0 — the narrative basis is carried in
# rp_note rather than replacing the number.
fmt_criterion <- function(value, criterion_form = NULL) {
  ifelse(is.na(value), "--", as.character(value))
}

# "Calculated" = TSD multiplying-factor projection. "Observed" = the value as
# reported, with no MF and no dilution ratio applied. Distinguishing these in
# the report matters: an observed DO minimum must not read as dilution-adjusted.
basis_label <- function(rp_method, rp_operator = NULL) {
  m <- ifelse(is.na(rp_method), "projected", rp_method)
  op <- if (is.null(rp_operator)) rep(NA_character_, length(m)) else rp_operator
  ifelse(
    m == "direct",
    ifelse(!is.na(op) & op == "<", "Observed min", "Observed max"),
    "Calculated"
  )
}

# wqs_info is now the full crosswalk_effective — all columns preserved.
# Each section that uses it selects the columns it needs locally.
wqs_info_df <- if (nzchar(params$wqs_info_path)) {
  readr::read_csv(params$wqs_info_path, show_col_types = FALSE) %>%
    dplyr::mutate(CRITERION_ID = as.character(CRITERION_ID))
} else NULL

overrides_df <- if (nzchar(params$overrides_path)) {
  readr::read_csv(params$overrides_path, show_col_types = FALSE)
} else NULL

# Water classes the analysis was run against, for the summary header. Taken
# from the RP results so manually added classes are included.
water_classes_txt <- rp_df %>%
  dplyr::filter(!is.na(USE_CLASS_NAME_LOCATION_ETC)) %>%
  dplyr::mutate(wc = format_water_class(
    USE_CLASS_NAME_LOCATION_ETC,
    if ("sub_class" %in% names(.)) sub_class else NA_character_
  )) %>%
  dplyr::distinct(wc) %>%
  dplyr::arrange(wc) %>%
  dplyr::pull(wc) %>%
  paste(collapse = ", ")

quick_run       <- isTRUE(params$quick_run)
quick_run_params_vec <- if (nzchar(params$quick_run_params %||% "")) {
  trimws(strsplit(params$quick_run_params, ",")[[1]])
} else character(0)

# Count outfalls (unique perm_feature_nmbr)
outfall_count <- if (!is.null(dmr_df) && "perm_feature_nmbr" %in% names(dmr_df)) {
  dplyr::n_distinct(dmr_df$perm_feature_nmbr)
} else {
  0L
}

# Permit limit history — guarded for standalone runs where ECHO limit columns may be absent
# perm_feature_nmbr is included so the table can be broken down by outfall.
limits_required_cols <- c("parameter_code", "limit_value_nmbr", "limit_unit_desc",
                           "limit_begin_date", "limit_end_date")
# Mass-load unit patterns — these are effluent load (mass/time) measurements that
# cannot be compared to concentration-based WQS criteria and are reported separately.
mass_load_units <- c("kg/d", "kg/day", "lbs/d", "lb/d", "lbs/day", "lb/day",
                     "mg/d", "mg/day", "g/d", "g/day", "ton/d", "ton/day",
                     "kg/mo", "lbs/mo", "kg/yr", "lbs/yr")

limits <- if (!is.null(dmr_df) && all(limits_required_cols %in% names(dmr_df))) {
  base <- dmr_df %>%
    dplyr::select(
      dplyr::all_of(limits_required_cols),
      dplyr::any_of("perm_feature_nmbr")   # include outfall if present
    ) %>%
    tidyr::drop_na(dplyr::all_of(limits_required_cols)) %>%
    dplyr::mutate(
      limit_begin_date = suppressWarnings(dplyr::coalesce(lubridate::ymd(limit_begin_date), lubridate::mdy(limit_begin_date))),
      limit_end_date   = suppressWarnings(dplyr::coalesce(lubridate::ymd(limit_end_date),   lubridate::mdy(limit_end_date))),
      is_mass_load     = tolower(trimws(limit_unit_desc)) %in%
                           tolower(mass_load_units)
    ) %>%
    dplyr::distinct()
  base
} else {
  NULL
}

# Metal criterion IDs and constants (for on-the-fly hardness curves in the report)
metals_ids <- c("79228","79238","79236","79253","79248","79251","79264")

metal_constants <- data.frame(
  CRITERION_ID = metals_ids,
  LC = c(0.7977, 0.8190, 0.8545, 1.273, 0.8460, 1.72, 0.8473),
  RC = c(3.909, 0.6848, 1.702, 4.705, 0.0584, 6.59, 0.884)
)

# Build curve of limits for a given CRITERION_ID across hardness 25..400 (step = 1)
metal_limit_curve <- function(crit_id, h_min = 25, h_max = 400) {
  row <- metal_constants[metal_constants$CRITERION_ID == crit_id, ]
  if (nrow(row) == 0) return(NULL)
  hardness <- seq(h_min, h_max, 1)
  limit <- exp(row$LC * log(hardness) - row$RC)
  dplyr::tibble(CRITERION_ID = crit_id, hardness = hardness, limit = limit)
}

# ── Shared visual palette ────────────────────────────────────────────────────
# Single source of truth for water class colors and linetypes.
# Used identically in pH, temperature, and concentration-based plots so the
# legend is immediately recognisable across all sections of the report.
#
# Drawing order rule: WQS lines are drawn thickest → thinnest so no line is
# buried under a wider one.  Within WQS lines, Min bounds (lwr_lwd) are drawn
# before Max bounds (upr_lwd); the RWC line is drawn last (on top).
#
wqs_colors <- c(
  "class SB waters"                 = "#1f78b4",  # blue
  "class SD waters"                 = "#ff7f00",  # orange
  "class SD waters- drinking water" = "#33a02c",  # green
  "class SG waters"                 = "#6a3d9a",  # purple
  "class SG waters- drinking water" = "#e31a1c",  # red
  "surface waters"                  = "#b15928",  # brown
  # Sub-class variants (PR class SD: stream vs reservoir/lake). Keyed by the
  # same combined label format_water_class() produces in the table cells, so
  # legends across the report stay consistent.
  "class SD waters \u2014 stream"         = "#cc6600",  # darker orange
  "class SD waters \u2014 reservoir/lake" = "#ffaa55"   # lighter orange
)

wqs_linetypes <- c(
  "class SB waters"                 = "dashed",
  "class SD waters"                 = "dashed",
  "class SD waters- drinking water" = "dotdash",
  "class SG waters"                 = "dashed",
  "class SG waters- drinking water" = "dotdash",
  "surface waters"                  = "longdash",
  "class SD waters \u2014 stream"         = "dashed",
  "class SD waters \u2014 reservoir/lake" = "dashed"
)

# Line widths: Max-bound lines are thinner and drawn on top; Min-bound lines
# are thicker and drawn first so Max is always visible above them.
wqs_lwd_max <- 0.7   # upper limit lines (Max pH, or WQS criterion for conc.)
wqs_lwd_min <- 1.3   # lower limit lines (Min pH)
rwc_color   <- "#2c7fb8"
rwc_lwd     <- 1.0   # RWC line — drawn last, distinct colour keeps it visible
limit_color <- "#d61309"  # hardness curve line colour

# Preset point colors by reported unit (dmr_unit_desc) — unchanged from before
unit_levels <- c(
  "deg C","mg/L","SU","mL/L","ug/L","MGD","NTU","col unit (pc)",
  "MF/L","kg/d","#/100mL","%","d","gal/d","MPN/100mL","m3/d"
)
unit_colors <- setNames(
  c(
    "#1b9e77", "#d95f02", "#7570b3", "#e7298a",
    "#66a61e", "#e6ab02", "#a6761d", "#1f78b4",
    "#b2df8a", "#fb9a99", "#e31a1c", "#ff7f00",
    "#6a3d9a", "#b15928", "#33a02c", "#a6cee3"
  ),
  unit_levels
)

# ── Shared per-parameter helpers ─────────────────────────────────────────────
# Used by both the Non-Concentration and Concentration sections so the two
# share one table/plot layout.

latex_escape <- function(x) gsub("_", "\\\\_", x)

outfall_type <- function(of) {
  if (!"perm_feature_type_code" %in% names(rp_df) || is.na(of)) return(NA_character_)
  rp_df %>%
    dplyr::filter(perm_feature_nmbr == of) %>%
    dplyr::pull(perm_feature_type_code) %>%
    dplyr::first()
}

# Dated DMR observations for one parameter at one outfall, optionally limited
# to the statistics a criterion compares (e.g. pH DAILY_MAX + DAILY_MIN).
dmr_obs <- function(codes, of, stats = NULL) {
  if (is.null(dmr_df)) return(data.frame())
  out <- dmr_df %>% dplyr::filter(parameter_code %in% codes)
  if (!is.na(of) && "perm_feature_nmbr" %in% names(out)) {
    out <- out %>% dplyr::filter(perm_feature_nmbr == of)
  }
  if (!is.null(stats) && "statistic" %in% names(out)) {
    out <- out %>% dplyr::filter(statistic %in% stats)
  }
  out %>%
    dplyr::mutate(
      # Handle both ISO YYYY-MM-DD (SQLite) and MM/DD/YYYY (ECHO legacy)
      monitoring_period_end_date = suppressWarnings(dplyr::coalesce(
        lubridate::ymd(monitoring_period_end_date),
        lubridate::mdy(monitoring_period_end_date)
      ))
    ) %>%
    dplyr::filter(!is.na(monitoring_period_end_date), !is.na(dmr_value_nmbr)) %>%
    dplyr::arrange(monitoring_period_end_date)
}

units_converted_label <- function(obs) {
  if (!"Conv_Flag" %in% names(obs) || nrow(obs) == 0) return("--")
  if (any(obs$Conv_Flag == "PASS", na.rm = TRUE)) "Yes" else "No"
}

manual_wqs_callout <- function(codes) {
  if (is.null(overrides_df) || all(is.na(codes))) return(invisible())
  ov <- overrides_df %>%
    dplyr::filter(parameter_code %in% codes, decision == "include")
  if (nrow(ov) > 0) {
    cat("> **\\textcolor{orange}{Manual WQS Applied}** — ",
        "Criterion value manually entered by user. ",
        "See *Manual WQS Overrides* section for full details.\n\n",
        sep = "")
  }
}

# Palette lookups with a fallback, so a class missing from the palette (e.g.
# one entered manually) still draws a line instead of being silently dropped.
class_colors <- function(classes) {
  classes <- unique(classes)
  setNames(dplyr::coalesce(unname(wqs_colors[classes]), "grey40"), classes)
}
class_linetypes <- function(classes) {
  classes <- unique(classes)
  setNames(dplyr::coalesce(unname(wqs_linetypes[classes]), "dashed"), classes)
}

# Temperature reported under 00011 (deg F) is converted to deg C upstream
# without checking the values. A permittee that enters Celsius numbers under
# the Fahrenheit code ends up with sub-zero results. The values are left as
# converted; a reading below freezing is flagged for the permit writer.
negative_temp_flag <- function(obs) {
  t_obs <- obs %>% dplyr::filter(parameter_code == "00010")
  n_neg <- sum(t_obs$dmr_value_nmbr < 0, na.rm = TRUE)
  if (n_neg == 0) return(invisible())
  from_f <- "orig_parameter_code" %in% names(t_obs) &&
    any(t_obs$orig_parameter_code[t_obs$dmr_value_nmbr < 0] == "00011", na.rm = TRUE)
  cat(sprintf(paste0(
    "> **\\textcolor{red}{Check data}** — %d of %d temperature values at this ",
    "outfall are below 0 °C%s. Verify against the DMR before relying on this result.\n\n"),
    n_neg, nrow(t_obs),
    if (from_f) paste0(
      ". They were reported under parameter 00011 (°F) and converted to °C; ",
      "values below freezing suggest they were actually reported in °C"
    ) else ""
  ))
}

obs_point_color <- "#1b9e77"

# Plot title naming the parameter and outfall, so a plot still identifies
# itself when its section runs over several pages. Internal Outfalls in red.
plot_title <- function(param, of) {
  type <- outfall_type(of)
  lbl <- format_outfall_label(of, type)
  if (!is.na(type) && type == "INO") lbl <- paste0("<span style='color:red'>", lbl, "</span>")
  paste0(param, " — ", lbl)
}

# Start each parameter on a new page unless its heading, note and first
# table+plot block fit, so a heading is never stranded at a page bottom.
# Skipped for a section's first parameter, which follows the section's own
# \clearpage and short intro: requiring the space there pushed it to the next
# page and left the intro alone on a page.
start_parameter_section <- function(first = FALSE) {
  if (first) return(invisible())
  cat("\n\n```{=latex}\n\\needspace{0.7\\textheight}\n```\n\n")
}

# A parameter's summary table and plot go in one minipage, which LaTeX cannot
# split: if the pair doesn't fit, it moves to the next page whole. The former
# samepage groups did not hold a table and figure together, so the plot moved
# alone and left the table stranded above a large blank gap.
keep_together_begin <- function() {
  cat("\n\n```{=latex}\n\\noindent\\begin{minipage}{\\linewidth}\n```\n\n")
}
keep_together_end <- function() {
  cat("\n\n```{=latex}\n\\end{minipage}\n\\par\\medskip\n```\n\n")
}

# Every plot gets a frame so its legend and caption read as part of the
# figure rather than running into the surrounding text.
report_plot_theme <- ggplot2::theme_minimal() +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(colour = "grey40", fill = NA, linewidth = 0.6),
    plot.margin     = ggplot2::margin(8, 10, 8, 8),
    plot.title      = ggtext::element_markdown(size = 10, face = "bold"),
    legend.position = "bottom",
    legend.box      = "vertical",
    legend.text     = ggplot2::element_text(size = 7),
    legend.title    = ggplot2::element_text(size = 8),
    plot.caption    = ggplot2::element_text(hjust = 0, color = "grey30", size = 8)
  )

# Observations against criterion lines for a direct-comparison parameter.
# Floors (pH minimum) are drawn thicker and first so the ceiling line on top
# stays visible when the two coincide.
plot_direct <- function(obs, rows, title) {
  wqs <- rows %>%
    dplyr::mutate(
      line_key = format_water_class(
        USE_CLASS_NAME_LOCATION_ETC,
        if ("sub_class" %in% names(.)) sub_class else NA_character_
      ),
      is_floor = !is.na(rp_operator) & rp_operator == "<",
      lwd = dplyr::if_else(is_floor, wqs_lwd_min, wqs_lwd_max)
    ) %>%
    dplyr::distinct(line_key, CRITERION_VALUE, is_floor, lwd) %>%
    dplyr::filter(is.finite(CRITERION_VALUE)) %>%
    dplyr::arrange(dplyr::desc(is_floor))

  pos  <- obs$dmr_value_nmbr[is.finite(obs$dmr_value_nmbr) & obs$dmr_value_nmbr > 0]
  crit <- suppressWarnings(min(wqs$CRITERION_VALUE[wqs$CRITERION_VALUE > 0]))
  use_log <- length(pos) > 0 && is.finite(crit) && crit > 0 && max(pos) / crit >= 100

  # Non-detects arrive as NODI B/Q rewritten to 0; log10 would drop them, so
  # they are plotted at the smallest positive value as open down-triangles.
  # Only exact zeros: a negative here is a real (if suspect) temperature and
  # must plot as a measurement.
  obs$is_nd <- is.finite(obs$dmr_value_nmbr) & obs$dmr_value_nmbr == 0
  nd_floor <- if (length(pos) > 0) min(pos) else NA_real_
  nd_at_floor <- use_log && any(obs$is_nd) && is.finite(nd_floor)
  obs$y_plot <- ifelse(obs$is_nd & nd_at_floor, nd_floor, obs$dmr_value_nmbr)
  nd_cap <- if (nd_at_floor) {
    paste0("Open triangles are non-detects, plotted at ", signif(nd_floor, 3), ".")
  } else if (any(obs$is_nd)) {
    "Open triangles are non-detects."
  } else {
    NULL
  }

  unit <- rows$UNIT_NAME[!is.na(rows$UNIT_NAME)][1]
  p <- ggplot2::ggplot(obs, ggplot2::aes(x = monitoring_period_end_date, y = y_plot)) +
    ggplot2::geom_line(color = "lightgrey", linetype = "dotted", linewidth = 0.5)

  if (nrow(wqs) > 0) {
    p <- p +
      ggplot2::geom_hline(
        data = wqs,
        ggplot2::aes(yintercept = CRITERION_VALUE, color = line_key,
                     linetype = line_key, linewidth = lwd)
      ) +
      ggplot2::scale_linewidth_identity() +
      ggplot2::scale_color_manual(name = "Water Class", values = class_colors(wqs$line_key)) +
      ggplot2::scale_linetype_manual(name = "Water Class", values = class_linetypes(wqs$line_key))
  }

  p <- p +
    ggplot2::geom_point(data = obs[!obs$is_nd, , drop = FALSE],
                        color = obs_point_color, alpha = 0.85)
  if (any(obs$is_nd)) {
    p <- p + ggplot2::geom_point(data = obs[obs$is_nd, , drop = FALSE],
                                 shape = 6, color = "grey30", size = 2.2, stroke = 0.9)
  }

  p +
    (if (use_log) ggplot2::scale_y_log10() else ggplot2::scale_y_continuous()) +
    ggplot2::labs(
      title = title,
      x = "Monitoring date",
      y = paste0(if (is.na(unit)) "" else unit, if (use_log) " (log scale)" else ""),
      caption = nd_cap
    ) +
    report_plot_theme
}
```

```{r}
#| label: title-block
#| results: asis

# Escape LaTeX specials in user/ECHO-supplied text (facility names can carry
# "&", "#", "%").
tex_escape <- function(x) {
  x <- gsub("\\", "\\textbackslash{}", x, fixed = TRUE)
  x <- gsub("([&%$#_{}])", "\\\\\\1", x)
  x <- gsub("~", "\\textasciitilde{}", x, fixed = TRUE)
  gsub("^", "\\textasciicircum{}", x, fixed = TRUE)
}
permit_tex   <- tex_escape(params$permit_id %||% "")
facility_tex <- tex_escape(params$facility_name %||% "")
generated    <- format(Sys.Date(), "%B %d, %Y")

detail_row <- function(label, value) {
  if (!nzchar(value)) return("")
  sprintf("\\textsf{\\textbf{\\textcolor{epablue}{%s}}} & %s \\\\[3pt]\n", label, value)
}

cat("```{=latex}\n")
cat(sprintf("\\gdef\\rppermit{%s}\n\\gdef\\rpgenerated{%s}\n",
            if (nzchar(permit_tex)) paste("Permit", permit_tex) else "", generated))
cat("\\thispagestyle{rptitle}\n")
# Logo beside the title, on white (the logo's own white disc would merge
# into a coloured band).
cat("\\noindent\\begin{minipage}[c]{1.25in}\\rplogo{1.1in}\\end{minipage}%\n")
cat("\\begin{minipage}[c]{\\dimexpr\\linewidth-1.25in\\relax}\\sffamily\n")
cat("{\\normalsize\\bfseries\\color{inkgray}U.S. Environmental Protection Agency \\textbar{} Region 2}\\\\[2pt]\n")
cat("{\\small\\color{inkgray}National Pollutant Discharge Elimination System}\\\\[4pt]\n")
cat("{\\color{epablue}\\LARGE\\bfseries Reasonable Potential Analysis Report}\n")
cat("\\end{minipage}\n\n\\vspace{16pt}\n")
# Permit band: white on EPA blue, which prints as white on dark grey.
cat("\\noindent\\colorbox{epablue}{\\parbox{\\dimexpr\\linewidth-2\\fboxsep\\relax}{\\vspace{7pt}\\sffamily\\color{white}\\hspace{6pt}")
cat(sprintf("{\\Large\\bfseries %s}", if (nzchar(permit_tex)) paste("Permit", permit_tex) else "NPDES Permit"))
if (nzchar(facility_tex)) cat(sprintf("\\\\[3pt]\\hspace*{6pt}{\\large %s}", facility_tex))
cat("\\vspace{7pt}}}\n\n")
cat("\\noindent{\\color{epagreen}\\rule{\\linewidth}{2.5pt}}\n\n\\vspace{10pt}\n")
cat("\\noindent\\begin{tabular}{@{}l l@{}}\n")
cat(detail_row("Monitoring period",
               if (nzchar(params$date_start)) paste(params$date_start, "to", params$date_end) else ""))
cat(detail_row("Application forms", tex_escape(params$forms %||% "")))
cat(detail_row("Report generated", generated))
cat("\\end{tabular}\n\n\\vspace{18pt}\n")
cat("\\setcounter{tocdepth}{1}\n{\\hypersetup{linkcolor=inkgray}\\tableofcontents}\n")
cat("```\n")
```

<!-- Each top-level section starts on a new page. -->
```{=latex}
\clearpage
```

# Summary

Permit: `r params$permit_id`  
Facility: `r params$facility_name`  
Date range: `r params$date_start` to `r params$date_end`  
Water class(es): `r water_classes_txt`  
Selected forms: `r params$forms`  
Dilution ratio: `r params$dilution_ratio`  
Confidence level: `r params$confidence_level`  
Target percentile: `r params$target_percentile`  
Hardness setting: `r if (params$hardness_mode == "set") paste0("Set (", params$hardness_value, " mg/L as CaCO3)") else "Range"`  

```{r}
#| label: quick-run-warning
#| echo: false
#| results: asis
if (isTRUE(quick_run) && length(quick_run_params_vec) > 0) {
  cat("\n\n> **\\textcolor{red}{WARNING — PRELIMINARY REPORT (Quick Run)}**\n>\n")
  cat("> This report was generated using Quick Run. The following parameters had unresolved\n")
  cat("> data issues and were **automatically excluded** from analysis. Results should be\n")
  cat("> considered preliminary. Re-run with all parameters reviewed for a final report.\n>\n")
  cat(paste0("> *Excluded parameters: ", paste(quick_run_params_vec, collapse = "; "), "*\n\n"))
}
```

## Included Pollutants

The following table shows the available data for each pollutant included in this analysis. The *Basis for Inclusion* column records why each pollutant is present: required by a selected NPDES form (`r params$forms`); carried in because prior DMR data exists even though no selected form requires it (*Existing DMR*); or present only because the user supplied the monitoring data (*Manually Added*). A \ddag\ marker flags parameters whose record includes monitoring data manually added by the user on top of the base DMR record. For a complete list of pollutants that had no associated data, refer to @sec-unavailable.

```{r}
#| label: coverage
#| echo: false

if (!is.null(coverage_df) && nrow(coverage_df) > 0) {

  # Determine which parameter codes had manual WQS entries (for dagger marker)
  manual_codes <- if (!is.null(overrides_df) && nrow(overrides_df) > 0) {
    overrides_df %>%
      dplyr::filter(decision == "include") %>%
      dplyr::pull(parameter_code)
  } else character(0)

  # `User Data` (from coverage.csv): "" none / "partial" some rows
  # user-supplied / "all" every row user-supplied. Older coverage files
  # without the column degrade gracefully to "".
  has_user_col <- "User Data" %in% names(coverage_df)

  top <- coverage_df %>%
    dplyr::filter(`# Samples` > 0) %>%
    dplyr::arrange(Pollutant) %>%
    dplyr::distinct() %>%
    dplyr::mutate(
      # as.character() guards the readr type guess: a coverage.csv where
      # every User Data value is empty parses as an all-NA *logical* column,
      # and coalesce(<logical>, "") is a type error that kills the chunk.
      user_flag = if (has_user_col) {
        dplyr::coalesce(as.character(`User Data`), "")
      } else {
        ""
      },
      # Basis for Inclusion:
      #   * form-required            -> "Form 2C" / "Forms 2C & 2F"
      #   * non-form, base DMR data  -> "Existing DMR"
      #   * non-form, all user rows  -> "Manually Added"
      `Basis for Inclusion` = dplyr::case_when(
        Form == "Not in selected forms" & user_flag == "all" ~ "Manually Added",
        Form == "Not in selected forms" ~ "Existing DMR",
        grepl("[,&]", Form) ~ paste("Forms", Form),
        TRUE ~ paste("Form", Form)
      ),
      # Two independent footnote markers:
      #   dagger  — WQS criterion manually entered (unchanged)
      #   ddagger — record includes user-added monitoring rows, except when
      #             the basis already says "Manually Added" (redundant there)
      Flag = paste0(
        dplyr::if_else(`Ref Param` %in% manual_codes, "\u2020", ""),
        dplyr::if_else(
          user_flag != "" & `Basis for Inclusion` != "Manually Added",
          "\u2021",
          ""
        )
      )
    ) %>%
    # No sample count here: coverage counts the whole facility, while every
    # other count in the report is per outfall.
    dplyr::select(
      Pollutant, `Basis for Inclusion`, `Ref Param`, Flag
    )

  tbl <- kable(top,
      format    = "latex",
      booktabs  = TRUE,
      longtable = TRUE,
      col.names = c("Pollutant", "Basis for Inclusion", "Ref Param", "")) %>%
    kable_styling(latex_options = c("striped", "repeat_header"))

  # Print each footnote only when its marker actually appears in a cell;
  # an unconditional footnote implies flagged rows that don't exist (and in
  # a run with no overrides and no user-added data, no footnote at all).
  has_dagger  <- any(grepl("\u2020", top$Flag))
  has_ddagger <- any(grepl("\u2021", top$Flag))
  fn_texts <- c(
    if (has_dagger) {
      "WQS criterion manually entered by user; see Manual WQS Overrides section."
    } else {
      NULL
    },
    if (has_ddagger) {
      "Includes monitoring data manually added by the user."
    } else {
      NULL
    }
  )
  fn_marks <- c(
    if (has_dagger) "\u2020" else NULL,
    if (has_ddagger) "\u2021" else NULL
  )
  if (length(fn_texts) > 0) {
    tbl <- tbl %>%
      footnote(symbol = fn_texts, symbol_manual = fn_marks, escape = FALSE)
  }
  tbl

} else {
  cat("No coverage summary available.")
}
```

```{=latex}
\clearpage
```

# RP Findings

RP column: **YES** indicates the assessed value exceeds the criterion.
**DEPENDS** indicates an exceedance was not observed but the tool cannot
conclude no reasonable potential — see the note on the pollutant's section.
**NO** indicates no exceedance.

The Basis column distinguishes a *Calculated* receiving water concentration
(TSD multiplying-factor projection, dilution ratio applied) from an
*Observed* maximum or minimum compared directly to the criterion with no
projection and no dilution. Values rounded to 2 decimals.

```{r}
#| label: rp-table
#| echo: false

rp_df %>%
  mutate(Basis = basis_label(
           if ("rp_method" %in% names(.)) rp_method else NA_character_,
           if ("rp_operator" %in% names(.)) rp_operator else NULL
         ),
         RWC_rs = round(RWC_rs, 2),
         CRITERION_VALUE = round(CRITERION_VALUE,2),
         CRITERION_ID = gsub("_", "\\\\_", CRITERION_ID),
         NPDES_Pollutant = gsub("_", "\\\\_", NPDES_Pollutant)) %>%
  drop_na(RWC_rs) %>%
  mutate(USE_CLASS_NAME_LOCATION_ETC = format_water_class(
           USE_CLASS_NAME_LOCATION_ETC,
           if ("sub_class" %in% names(.)) sub_class else NA_character_
         ),
         # Internal Outfalls: red bold text via cell_spec so they stay
         # visibly distinct from External Outfalls in the PDF.
         perm_feature_nmbr = {
           lbl <- format_outfall_label(
             perm_feature_nmbr,
             if ("perm_feature_type_code" %in% names(.)) perm_feature_type_code else NA_character_
           )
           is_int <- !is.na(perm_feature_type_code) & perm_feature_type_code == "INO"
           ifelse(
             is_int,
             kableExtra::cell_spec(lbl, format = "latex", color = "red", bold = TRUE),
             lbl
           )
         }) %>%
  select(CRITERION_ID, NPDES_Pollutant, perm_feature_nmbr, USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE, RWC_rs, Basis, RP) %>%
  # Severity order YES > DEPENDS > NO. A plain desc() on the character column
  # would sort YES, NO, DEPENDS and bury the determinations needing review.
  arrange(desc(factor(RP, levels = c("NO", "DEPENDS", "YES"), ordered = TRUE)),
          NPDES_Pollutant) %>%
  distinct() %>%
  # \mbox stops LaTeX hyphenating "DEPENDS" in the narrow RP column.
  mutate(RP = ifelse(is.na(RP), RP, paste0("\\mbox{", RP, "}"))) %>%
  kable(
    col.names = c("WQS ID","Pollutant","Outfall","Water Class","WQS Criterion","Value","Basis","RP"),
    format    = "latex",
    booktabs  = TRUE,
    longtable = TRUE,
    escape    = FALSE
  ) %>%
  kable_styling(latex_options = c("striped", "repeat_header"),
                font_size = 9) %>%
  # Explicit widths so all 8 columns fit within the 6.5in text block.
  # Sum = 14.3cm, leaving room for inter-column padding. WQS ID is wide
  # enough for manual IDs ("MANUAL-00010"), Basis for "Calculated" and RP
  # for "DEPENDS" without hyphenation.
  column_spec(1, width = "1.9cm") %>%   # WQS ID
  column_spec(2, width = "2.5cm") %>%   # Pollutant
  column_spec(3, width = "1.9cm") %>%   # Outfall (wraps for plain + coloured labels)
  column_spec(4, width = "2.1cm") %>%   # Water Class
  column_spec(5, width = "1.4cm") %>%   # WQS Criterion
  column_spec(6, width = "1.3cm") %>%   # Value (calculated RWC or observed)
  column_spec(7, width = "1.7cm") %>%   # Basis
  column_spec(8, width = "1.5cm")       # RP
```

```{=latex}
\clearpage
```

# Non-Concentration Based Measures

Parameters in this section are assessed by direct comparison: the observed
maximum (for a ceiling criterion) or minimum (for a floor) is compared to the
criterion as reported. No multiplying factor is applied and the dilution ratio
is not used.

```{r}
#| label: receiving-water
#| echo: false
#| fig-height: 2.9
#| results: asis

# Short enough that two table+plot blocks fit on one page. Set here because
# knitr ignores opts_current$set() for plots printed inside a loop.

direct_rp <- if ("rp_method" %in% names(rp_df)) {
  rp_df %>%
    dplyr::filter(!is.na(rp_method), rp_method == "direct") %>%
    dplyr::mutate(
      # pH carries two criterion rows (Max / Min) that belong in one section
      # with one table; every other direct parameter is keyed by its display
      # name, which keeps the two Enterococci statistics apart.
      section = dplyr::if_else(
        parameter_code == "00400",
        "pH",
        if ("criterion_label" %in% names(.)) {
          dplyr::coalesce(dplyr::na_if(criterion_label, ""), NPDES_Pollutant)
        } else {
          NPDES_Pollutant
        }
      )
    )
} else {
  rp_df[0, ]
}

if (nrow(direct_rp) == 0) {
  cat("No direct-comparison parameters were assessed.\n\n")
} else {
  for (sec in sort(unique(direct_rp$section))) {
    sec_rows <- direct_rp %>% dplyr::filter(section == sec)
    sec_codes <- unique(sec_rows$parameter_code)

    start_parameter_section(first = sec == sort(unique(direct_rp$section))[1])
    cat(sprintf("\n\n## %s\n\n", latex_escape(sec)))

    # Criterion note (Streeter-Phelps, Enterococci interval, temperature basis)
    if ("rp_note" %in% names(sec_rows)) {
      nt <- unique(sec_rows$rp_note)
      nt <- nt[!is.na(nt) & nzchar(nt)]
      if (length(nt) > 0) {
        cat(sprintf("> **\\textcolor{orange}{Note}** — %s\n\n", latex_escape(nt[1])))
      }
    }
    manual_wqs_callout(sec_codes)

    ops <- unique(stats::na.omit(sec_rows$rp_operator))
    cmp_txt <- if (all(c("<", ">") %in% ops)) {
      "The observed maximum is compared to the upper criterion and the observed minimum to the lower criterion"
    } else if (identical(ops, "<")) {
      "The observed minimum is compared to the criterion"
    } else {
      "The observed maximum is compared to the criterion"
    }
    cat(sprintf(
      "This parameter is assessed by direct comparison. %s as reported: no multiplying factor is applied and the dilution ratio is not used.\n\n",
      cmp_txt
    ))

    two_bound <- all(c("<", ">") %in% ops)

    for (of in sort(unique(stats::na.omit(sec_rows$perm_feature_nmbr)))) {
      of_rows <- sec_rows %>% dplyr::filter(perm_feature_nmbr == of)
      obs <- dmr_obs(sec_codes, of, unique(of_rows$statistic))

      cat(sprintf("### %s\n\n", format_outfall_label_latex(of, outfall_type(of))))
      negative_temp_flag(obs)
      keep_together_begin()

      summ <- of_rows %>%
        dplyr::mutate(
          water = format_water_class(
            USE_CLASS_NAME_LOCATION_ETC,
            if ("sub_class" %in% names(.)) sub_class else NA_character_
          ),
          bound = dplyr::if_else(!is.na(rp_operator) & rp_operator == "<", "Min", "Max"),
          samples = n_used,
          conv = units_converted_label(obs),
          crit = round(CRITERION_VALUE, 4),
          observed = round(
            if ("observed_value" %in% names(.)) dplyr::coalesce(observed_value, RWC_rs) else RWC_rs,
            4
          )
        ) %>%
        dplyr::arrange(water, dplyr::desc(bound)) %>%
        dplyr::select(CRITERION_ID, water, bound, samples, conv, crit, observed, RP) %>%
        dplyr::distinct()

      obs_hdr <- if (two_bound) "Observed" else if (identical(ops, "<")) "Observed Min" else "Observed Max"
      if (!two_bound) summ <- dplyr::select(summ, -bound)
      print(knitr::kable(
        summ, format = "latex", booktabs = TRUE,
        col.names = c("WQS ID", "Water Class", if (two_bound) "Bound",
                      "Samples", "Units Converted", "WQS Criterion", obs_hdr, "RP")
      # scale_down only shrinks a table wider than the text block (e.g. long
      # manual WQS IDs with three water classes); others print unchanged.
      ) %>% kableExtra::kable_styling(latex_options = c("HOLD_position", "scale_down")))
      cat("\n\n")

      if (nrow(obs) > 0) {
        print(plot_direct(obs, of_rows, plot_title(sec, of)))
        cat("\n\n")
      } else {
        cat("No dated observations to plot.\n\n")
      }
      keep_together_end()
    }
  }
}
```




```{=latex}
\clearpage
```

# Concentration Based Measures

## RP Equations

The following equations estimate the Receiving Water Concentration (RWC) for
parameters assessed by projection. Parameters assessed by direct comparison
are reported in the Non-Concentration Based Measures section.

$$\sigma_g = \sqrt{\ln(\mathrm{CV}^2 + 1)}$$
$$p = (1 - \mathrm{CL})^{1/n}$$
$$z_p = \Phi^{-1}(p),\; z_{\text{target}} = \Phi^{-1}(\mathrm{target})$$
$$\mathrm{MF} = \exp\big((z_{\text{target}} - z_p)\,\sigma_g\big)$$

$$\text{RWC} = \frac{\max(x) \times MF}{DR}$$

```{r}
#| label: per-pollutant
#| echo: false
#| fig-height: 2.9
#| results: asis

# Short enough that two table+plot blocks fit on one page. Set here because
# knitr ignores opts_current$set() for plots printed inside a loop.

if (is.null(dmr_df) || nrow(dmr_df) == 0) {
  cat("No DMR data were provided; per-pollutant sections are omitted.")
} else {
  pollutants <- rp_df %>%
    dplyr::filter(
      is.na(rp_method) | rp_method != "direct",
      (!is.na(n_used) & n_used > 0) | is.finite(RWC_rs)
    ) %>%
    dplyr::distinct(NPDES_Pollutant) %>%
    dplyr::pull()

  for (poll in pollutants) {
    sub_all <- rp_df %>%
      dplyr::filter(NPDES_Pollutant == poll, is.na(rp_method) | rp_method != "direct")
    param_codes <- sub_all %>% dplyr::distinct(parameter_code) %>% dplyr::pull()
    param_code <- if (length(param_codes) > 0) param_codes[1] else NA_character_

    start_parameter_section(first = poll == pollutants[1])
    cat(sprintf("\n\n## %s\n\n", latex_escape(poll)))

    # ── Criterion note (rp_note) ──────────────────────────────────────────
    if ("rp_note" %in% names(sub_all)) {
      note_txt <- sub_all %>%
        dplyr::pull(rp_note) %>%
        unique()
      note_txt <- note_txt[!is.na(note_txt) & nzchar(note_txt)]
      if (length(note_txt) > 0) {
        cat(sprintf(
          "> **\\textcolor{orange}{Note}** — %s\n\n",
          latex_escape(note_txt[1])
        ))
      }
    }
    manual_wqs_callout(param_code)

    outfalls_for_poll <- if ("perm_feature_nmbr" %in% names(sub_all)) {
      unique(sub_all$perm_feature_nmbr)
    } else if (!is.null(dmr_df) && "perm_feature_nmbr" %in% names(dmr_df)) {
      unique(dplyr::filter(dmr_df, parameter_code == param_code)$perm_feature_nmbr)
    } else NA_character_
    outfalls_for_poll <- outfalls_for_poll[!is.na(outfalls_for_poll)]
    if (length(outfalls_for_poll) == 0) outfalls_for_poll <- NA_character_

    show_outfall_subheads <- is.numeric(outfall_count) && outfall_count > 1

    for (of in sort(outfalls_for_poll)) {
      # \nopagebreak[4] strongly discourages a page break immediately before
      # the outfall subheading, keeping it with the first line of content.
      cat("\n\n\\nopagebreak[4]\n\n")

      if (show_outfall_subheads && !is.na(of)) {
        cat(sprintf("### %s\n\n", format_outfall_label_latex(of, outfall_type(of))))
      }

      sub_rp <- if (!is.na(of)) sub_all %>% dplyr::filter(perm_feature_nmbr == of) else sub_all
      obs <- dmr_obs(param_code, of)
      n_samples <- nrow(obs)

      rwc_val <- sub_rp %>%
        dplyr::distinct(RWC_rs) %>% dplyr::pull(RWC_rs) %>% .[is.finite(.)]
      rwc_val <- if (length(rwc_val) > 0) rwc_val[1] else NA_real_

      negative_temp_flag(obs)
      keep_together_begin()

      # ── Summary table ─────────────────────────────────────────────────────
      # The single table for the parameter at this outfall: one row per
      # criterion, so several water classes simply add rows.
      summ <- sub_rp %>%
        dplyr::mutate(
          water = format_water_class(
            USE_CLASS_NAME_LOCATION_ETC,
            if ("sub_class" %in% names(.)) sub_class else NA_character_
          ),
          samples = dplyr::coalesce(as.integer(n_used), n_samples),
          conv = units_converted_label(obs),
          rwc = signif(RWC_rs, 3)
        ) %>%
        dplyr::select(CRITERION_ID, water, samples, conv, CRITERION_VALUE, rwc, RP) %>%
        dplyr::distinct()

      if (nrow(summ) > 0) {
        kable(summ,
              format    = "latex",
              booktabs  = TRUE,
              col.names = c("WQS ID", "Water Class", "Samples", "Units Converted",
                            "WQS Criterion", "RWC", "RP")) %>%
          kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
          print()
        cat("\n\n")
      } else {
        cat("No WQS rows found for this pollutant/outfall.\n\n")
      }

      # wqs_lines drives the dashed criterion lines on the per-pollutant
      # plot AND keys their legend entries via wqs_colors / wqs_linetypes.
      # When a parameter has multiple criteria in the same base class that
      # differ by sub_class (e.g., phosphorus SD streams 160 ug/L vs SD
      # reservoir/lake 26 ug/L), we substitute the combined display label
      # so each criterion gets its own legend entry and its own palette
      # slot. Rows without sub_class fall through with the base class.
      wqs_lines <- sub_rp %>%
        dplyr::mutate(
          USE_CLASS_NAME_LOCATION_ETC = format_water_class(
            USE_CLASS_NAME_LOCATION_ETC,
            if ("sub_class" %in% names(.)) sub_class else NA_character_
          )
        ) %>%
        dplyr::select(CRITERION_ID, USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE) %>%
        dplyr::distinct() %>%
        dplyr::filter(is.finite(CRITERION_VALUE))

      # Hardness annotation (metals / range mode)
      hardness_note <- NULL
      if (!is.null(params$hardness_mode) &&
          params$hardness_mode == "range" && is.finite(rwc_val)) {
        sub_sd_metal <- sub_rp %>%
          dplyr::filter(CRITERION_ID %in% metals_ids,
                        USE_CLASS_NAME_LOCATION_ETC == "class SD waters")
        if (nrow(sub_sd_metal) > 0) {
          crit  <- sub_sd_metal$CRITERION_ID[1]
          curve <- metal_limit_curve(crit)
          if (!is.null(curve) && nrow(curve) > 0) {
            h_row <- curve %>%
              dplyr::mutate(diff = abs(limit - rwc_val)) %>%
              dplyr::slice_min(diff, n = 1, with_ties = FALSE)
            if (nrow(h_row) == 1 && is.finite(h_row$hardness)) {
              hardness_note <- paste0("RWC > WQS at Hardness values < ",
                                      round(h_row$hardness, 0),
                                      " mg/L as CaCO₃")
            }
          }
        }
      }

      # ── PLOT: Observed vs WQS / RWC ─────────────────────────────────────────
      # Drawing order: WQS lines (thicker, drawn first) → RWC line (on top).
      # WQS lines are drawn individually via a loop so each gets the correct
      # class colour and linetype from the shared palette.
      if (n_samples > 0 &&
          all(c("monitoring_period_end_date", "dmr_value_nmbr") %in% names(obs))) {

        ylab_txt         <- sub_rp$UNIT_NAME[1]
        present_classes  <- unique(wqs_lines$USE_CLASS_NAME_LOCATION_ETC)

        # y-axis bounds: include obs values, WQS lines, and RWC
        all_y_vals <- c(obs$dmr_value_nmbr, wqs_lines$CRITERION_VALUE,
                        if (is.finite(rwc_val)) rwc_val else numeric(0))
        # Floor at 0 unless the data go negative (a suspect temperature
        # conversion); clipping would hide exactly the values being flagged.
        y_lo  <- min(all_y_vals, na.rm = TRUE)
        y_min <- if (y_lo < 0) y_lo * 1.05 else y_lo * 0.95

        # Mirror the app: many CWA criteria sit far below any achievable
        # quantitation limit, and bacteria counts span five or six decades by
        # nature, so a linear axis collapses the criterion line onto zero.
        # Switch to log10 when that happens. This ratio is NOT a units check --
        # reported units are reconciled upstream by get_unit_conversion(), so a
        # large ratio here is normally a real exceedance or a naturally
        # wide-range parameter, not a mg/L <-> ug/L reporting error.
        pos_obs_r  <- obs$dmr_value_nmbr[is.finite(obs$dmr_value_nmbr) &
          obs$dmr_value_nmbr > 0]
        crit_min_r <- suppressWarnings(min(
          wqs_lines$CRITERION_VALUE[is.finite(wqs_lines$CRITERION_VALUE) &
            wqs_lines$CRITERION_VALUE > 0]
        ))
        oom_ratio_r <- if (length(pos_obs_r) > 0 && is.finite(crit_min_r) &&
          crit_min_r > 0) {
          max(pos_obs_r) / crit_min_r
        } else {
          NA_real_
        }
        use_log_r <- is.finite(oom_ratio_r) && oom_ratio_r >= 100
        if (use_log_r) ylab_txt <- paste0(ylab_txt, " (log scale)")

        # Non-detects (NODI B/Q, stored as 0) and literal reported zeros are
        # dropped silently by log10. Plot them at a floor equal to the smallest
        # positive value in the series, marked with open down-triangles, and
        # disclose the substitution in the caption.
        # Exact zeros only: a negative is a real (if suspect) temperature.
        obs$is_nd_r <- is.finite(obs$dmr_value_nmbr) & obs$dmr_value_nmbr == 0
        nd_floor_r <- if (length(pos_obs_r) > 0) min(pos_obs_r) else NA_real_
        nd_at_floor_r <- use_log_r && any(obs$is_nd_r) && is.finite(nd_floor_r)
        obs$y_plot_r <- ifelse(
          obs$is_nd_r & nd_at_floor_r,
          nd_floor_r,
          obs$dmr_value_nmbr
        )
        nd_cap_r <- if (nd_at_floor_r) {
          paste0("Open triangles are non-detects, plotted at ",
                 signif(nd_floor_r, 3), ".")
        } else if (any(obs$is_nd_r)) {
          "Open triangles are non-detects."
        } else {
          NA_character_
        }

        p <- ggplot2::ggplot(
          obs,
          ggplot2::aes(x = monitoring_period_end_date, y = y_plot_r)
        ) +
          ggplot2::geom_line(color = "lightgrey", linetype = "dotted",
                             linewidth = 0.5)

        # Draw WQS lines first (thicker), one at a time for full palette control
        if (nrow(wqs_lines) > 0) {
          for (i in seq_len(nrow(wqs_lines))) {
            row_i <- wqs_lines[i, ]
            p <- p + ggplot2::geom_hline(
              yintercept  = row_i$CRITERION_VALUE,
              color       = class_colors(row_i$USE_CLASS_NAME_LOCATION_ETC),
              linetype    = class_linetypes(row_i$USE_CLASS_NAME_LOCATION_ETC),
              linewidth   = wqs_lwd_max,
              inherit.aes = FALSE
            )
          }
        }

        # RWC line drawn last — on top of all WQS lines
        if (is.finite(rwc_val)) {
          p <- p + ggplot2::geom_hline(
            yintercept  = rwc_val,
            color       = rwc_color,
            linetype    = "solid",
            linewidth   = rwc_lwd,
            inherit.aes = FALSE
          )
        }

        # Observation points. The unit is already on the y-axis label, so the
        # points are coloured by reported unit (with a legend) only when the
        # record mixes units — the one case where the colour carries
        # information the axis label doesn't.
        mixed_units <- dplyr::n_distinct(obs$dmr_unit_desc[!obs$is_nd_r], na.rm = TRUE) > 1
        if (mixed_units) {
          p <- p +
            ggplot2::geom_point(
              data = obs[!obs$is_nd_r, , drop = FALSE],
              ggplot2::aes(color = dmr_unit_desc),
              alpha = 0.85
            ) +
            ggplot2::scale_color_manual(
              name     = "Reported Unit",
              values   = unit_colors,
              na.value = "#999999"
            )
        } else {
          p <- p + ggplot2::geom_point(
            data = obs[!obs$is_nd_r, , drop = FALSE],
            color = obs_point_color, alpha = 0.85
          )
        }

        p <- p +
          {
            if (use_log_r) {
              ggplot2::scale_y_log10()
            } else {
              ggplot2::scale_y_continuous(limits = c(y_min, NA))
            }
          } +
          ggplot2::labs(
            title = plot_title(poll, of),
            x = "Monitoring date",
            y = ylab_txt,
            caption = if (!is.na(nd_cap_r)) nd_cap_r else NULL
          ) +
          report_plot_theme +
          ggplot2::theme(
            plot.caption = ggplot2::element_text(hjust = 0, color = rwc_color, size = 8)
          )

        # Non-detect markers drawn after the unit scale so they take no colour
        # from it. inherit.aes keeps x and y_plot_r from the base mapping.
        if (any(obs$is_nd_r)) {
          p <- p + ggplot2::geom_point(
            data  = obs[obs$is_nd_r, , drop = FALSE],
            shape = 6, color = "grey30", size = 2.2, stroke = 0.9
          )
        }

        # Water class legend via a second colour scale (separate from units)
        if (length(present_classes) > 0) {
          p <- p + ggnewscale::new_scale_color() +
            ggplot2::geom_hline(
              data = wqs_lines,
              ggplot2::aes(yintercept  = CRITERION_VALUE,
                           color       = USE_CLASS_NAME_LOCATION_ETC,
                           linetype    = USE_CLASS_NAME_LOCATION_ETC),
              linewidth   = wqs_lwd_max,
              inherit.aes = FALSE,
              show.legend = TRUE
            ) +
            ggplot2::scale_color_manual(
              name   = "Water Class",
              values = class_colors(present_classes)
            ) +
            ggplot2::scale_linetype_manual(
              name   = "Water Class",
              values = class_linetypes(present_classes)
            )
        }

        # Hardness note goes in the caption (below the plot) rather than as an
        # in-panel annotation, which previously ran off the right page margin.
        if (!is.null(hardness_note) && is.finite(rwc_val)) {
          p <- p + ggplot2::labs(caption = hardness_note)
        }

        print(p)
        cat("\n\n")

      } else {
        cat("No dated observations to plot.\n\n")
      }
      keep_together_end()

      # ── Hardness vs Limit (metals / range mode only) ────────────────────────
      show_range_plot <-
        (!is.null(params$hardness_mode) && params$hardness_mode == "range") &&
        any(sub_rp$CRITERION_ID %in% metals_ids &
              sub_rp$USE_CLASS_NAME_LOCATION_ETC == "class SD waters")

      if (show_range_plot) {
        crit  <- sub_rp %>%
          dplyr::filter(CRITERION_ID %in% metals_ids,
                        USE_CLASS_NAME_LOCATION_ETC == "class SD waters") %>%
          dplyr::slice(1) %>% dplyr::pull(CRITERION_ID)
        curve <- metal_limit_curve(crit)
        if (!is.null(curve) && nrow(curve) > 0) {
          ylab_txt <- sub_rp$UNIT_NAME[1]
          p_h <- ggplot2::ggplot(curve, ggplot2::aes(x = hardness, y = limit)) +
            ggplot2::geom_line(color = limit_color, linewidth = 1) +
            ggplot2::scale_y_continuous(limits = c(0, NA)) +
            ggplot2::labs(
              title = paste("Hardness vs WQS Criterion —", poll, "(Class SD)"),
              x     = "Hardness (mg/L as CaCO₃)",
              y     = ylab_txt
            ) +
            report_plot_theme
          knitr::opts_current$set(fig.height = 3.2)
          print(p_h)
          cat("\n\n")
        }
      }

      # ── RWC calculation ─────────────────────────────────────────────────────
      cat("**RWC Calculation**\n\n")
      if (all(c("n_used","cv","sigma_g","p_percentile","z_p","z_target","MF") %in%
              names(sub_rp))) {
        n_val   <- sub_rp %>% dplyr::distinct(n_used)       %>% dplyr::pull() %>% .[is.finite(.)]
        cv_val  <- sub_rp %>% dplyr::distinct(cv)           %>% dplyr::pull() %>% .[is.finite(.)]
        sg_val  <- sub_rp %>% dplyr::distinct(sigma_g)      %>% dplyr::pull() %>% .[is.finite(.)]
        p_val   <- sub_rp %>% dplyr::distinct(p_percentile) %>% dplyr::pull() %>% .[is.finite(.)]
        zp_val  <- sub_rp %>% dplyr::distinct(z_p)          %>% dplyr::pull() %>% .[is.finite(.)]
        zt_val  <- sub_rp %>% dplyr::distinct(z_target)     %>% dplyr::pull() %>% .[is.finite(.)]
        mf_val  <- sub_rp %>% dplyr::distinct(MF)           %>% dplyr::pull() %>% .[is.finite(.)]
        rwc_eq  <- sub_rp %>% dplyr::distinct(RWC_rs)       %>% dplyr::pull() %>% .[is.finite(.)]
        max_val <- sub_rp %>% dplyr::distinct(max_value)    %>% dplyr::pull() %>% .[is.finite(.)]
        cl_val  <- params$confidence_level
        tgt_val <- params$target_percentile

        if (length(n_val) > 0 && length(cv_val) > 0 && length(sg_val) > 0 &&
            length(p_val) > 0 && length(zp_val) > 0 && length(zt_val) > 0 &&
            length(mf_val) > 0 && is.finite(cl_val) && is.finite(tgt_val)) {

          cat(sprintf("$$\\sigma_g = \\sqrt{\\ln(%s^2+1)} = %s$$\n",
                      format(signif(cv_val[1], 2), scientific = FALSE),
                      format(signif(sg_val[1], 2), scientific = FALSE)))
          cat(sprintf("$$p = (1 - %s)^{1/%s} = %s$$\n",
                      format(signif(cl_val, 2), scientific = FALSE),
                      format(n_val[1]),
                      format(signif(p_val[1], 2), scientific = FALSE)))
          cat(sprintf("$$z_p = %s,\\; z_{\\text{target}} = %s$$\n",
                      format(signif(zp_val[1], 2), scientific = FALSE),
                      format(signif(zt_val[1], 2), scientific = FALSE)))
          cat(sprintf(
            "$$\\mathrm{MF} = \\exp\\big((%s - %s)\\times %s\\big) = %s$$\n",
            format(signif(zt_val[1], 2), scientific = FALSE),
            format(signif(zp_val[1], 2), scientific = FALSE),
            format(signif(sg_val[1], 2), scientific = FALSE),
            format(signif(mf_val[1], 2), scientific = FALSE)))
          cat(sprintf(
            "$$\\mathrm{RWC} = \\frac{\\max(x)\\times MF}{DR} = \\frac{%s\\times%s}{%s} = %s$$\n\n",
            format(signif(max_val, 2), scientific = FALSE),
            format(signif(mf_val,  2), scientific = FALSE),
            format(signif(params$dilution_ratio, 2), scientific = FALSE),
            format(signif(rwc_eq,  2), scientific = FALSE)))
        }
      }

      # ── Permit Limit History (per outfall) ────────────────────────────────
      # Limits can vary by outfall, so the table is scoped to the current outfall.
      # Mass-load limits (e.g. kg/d) cannot be compared to concentration-based
      # WQS criteria and are listed separately.
      if (!is.null(limits) && is.data.frame(limits) && !is.na(param_code)) {
        lim_base <- limits %>%
          dplyr::filter(parameter_code == param_code)
        if ("perm_feature_nmbr" %in% names(lim_base) && !is.na(of)) {
          lim_base <- lim_base %>% dplyr::filter(perm_feature_nmbr == of)
        }

        lim_conc <- lim_base %>%
          dplyr::filter(!is_mass_load) %>%
          dplyr::transmute(
            `Effluent Limit` = limit_value_nmbr,
            `Unit`           = limit_unit_desc,
            `Limit Start`    = format(limit_begin_date, "%m/%d/%Y"),
            `Limit End`      = format(limit_end_date,   "%m/%d/%Y")
          ) %>%
          dplyr::distinct()

        if (nrow(lim_conc) > 0) {
          cat("**Permit Effluent Limit History**\n\n")
          kable(lim_conc,
                format    = "latex",
                booktabs  = TRUE,
                col.names = c("Effluent Limit", "Unit", "Limit Start", "Limit End")) %>%
            kable_styling(latex_options = c("HOLD_position")) %>%
            print()
          cat("\n\n")
        }

        lim_mass <- lim_base %>%
          dplyr::filter(is_mass_load) %>%
          dplyr::transmute(
            `Effluent Limit` = limit_value_nmbr,
            `Unit`           = limit_unit_desc,
            `Limit Start`    = format(limit_begin_date, "%m/%d/%Y"),
            `Limit End`      = format(limit_end_date,   "%m/%d/%Y")
          ) %>%
          dplyr::distinct()

        if (nrow(lim_mass) > 0) {
          cat("**Mass-Load Effluent Limits (not used in WQS analysis)**\n\n")
          cat(paste0("The following limits are expressed as mass-load quantities ",
                     "(mass per unit time) and cannot be directly compared to the ",
                     "concentration-based WQS criteria used in this analysis. ",
                     "They are provided for reference only.\n\n"))
          kable(lim_mass,
                format    = "latex",
                booktabs  = TRUE,
                col.names = c("Effluent Limit", "Unit", "Limit Start", "Limit End")) %>%
            kable_styling(latex_options = c("HOLD_position")) %>%
            print()
          cat("\n\n")
        }
      }

    } # end outfall loop
  } # end pollutant loop
} # end dmr_df check
```


```{=latex}
\clearpage
```

# Parameters from Applications without DMR Data {#sec-unavailable}

```{r}
#| label: unavailable
#| echo: false

# Rebuild bottom here so this chunk is self-contained and does not depend on
# the coverage chunk having run successfully (e.g. in standalone mode where
# coverage_df may be NULL).
bottom_unavail <- if (!is.null(coverage_df) && nrow(coverage_df) > 0) {
  coverage_df %>%
    dplyr::filter(`# Samples` == 0) %>%
    dplyr::arrange(Pollutant) %>%
    dplyr::select(!dplyr::any_of(c("Form", "User Data"))) %>%
    dplyr::distinct()
} else {
  NULL
}

if (!is.null(bottom_unavail) && nrow(bottom_unavail) > 0) {
  kable(bottom_unavail,
        format    = "latex",
        booktabs  = TRUE,
        longtable = TRUE,
        col.names = c("Pollutant", "Ref Param", "# Samples")) %>%
    kable_styling(latex_options = c("striped", "repeat_header"))
} else {
  cat("All expected pollutants had data available.")
}
```




```{=latex}
\clearpage
```

# Excluded Parameters {#sec-excluded}

```{r}
#| label: excluded-params
#| echo: false
#| results: asis

excl_df <- if (!is.null(overrides_df) && nrow(overrides_df) > 0) {
  overrides_df %>%
    dplyr::filter(decision == "exclude") %>%
    dplyr::select(
      `Parameter`      = parameter_desc,
      `Code`           = parameter_code,
      `Flag Reason`    = case_reason,
      `Samples in DMR` = dplyr::any_of("n_samples"),
      `Reason for Exclusion` = notes
    )
} else NULL

if (!is.null(excl_df) && nrow(excl_df) > 0) {
  cat("The following parameters had DMR data available but were excluded from the RP ",
      "analysis at the direction of the user.\n\n")
  tbl <- kable(excl_df,
        format    = "latex",
        booktabs  = TRUE,
        longtable = TRUE) %>%
    kable_styling(latex_options = c("striped", "repeat_header")) %>%
    # Widths on every column: with only Parameter and Reason fixed, a long
    # Flag Reason pushed the table past the right margin and the Reason for
    # Exclusion text was cut off. Content ~13.4cm + ~2.1cm padding fits 16.5cm.
    column_spec(1, width = "3.2cm") %>%   # Parameter
    column_spec(2, width = "1.2cm") %>%   # Code
    column_spec(3, width = "2.8cm")       # Flag Reason
  if (ncol(excl_df) == 5) tbl <- column_spec(tbl, 4, width = "1.4cm")  # Samples in DMR
  print(column_spec(tbl, ncol(excl_df), width = "4.8cm"))             # Reason for Exclusion
} else {
  cat("No parameters were explicitly excluded from this analysis.")
}
```


```{=latex}
\clearpage
```

# Manual WQS Overrides {#sec-manual-wqs}

```{r}
#| label: manual-wqs
#| echo: false
#| results: asis

manual_df <- if (!is.null(overrides_df) && nrow(overrides_df) > 0) {
  overrides_df %>%
    dplyr::filter(decision == "include") %>%
    dplyr::select(
      `Parameter`        = parameter_desc,
      `Code`             = parameter_code,
      `Flag Reason`      = case_reason,
      `Water Class(es)`  = water_classes,
      `Criterion Value`  = criterion_value,
      `Unit`             = unit,
      `Criterion Type`   = criterion_type,
      `Basis / Notes`    = notes
    )
} else NULL

if (!is.null(manual_df) && nrow(manual_df) > 0) {
  cat(paste0(
    "The following parameters had no matching WQS criterion for the selected water ",
    "class and were evaluated using manually entered criteria. These criteria were ",
    "supplied by the permit writer and are documented here as part of the permanent ",
    "record. All manually entered criteria are flagged with a dagger symbol in the Coverage ",
    "Summary table.\n\n"
  ))
  # print() is explicit: under results:asis an auto-printed kable can be
  # skipped, so force the LaTeX out.
  print(
  kable(manual_df,
        format    = "latex",
        booktabs  = TRUE,
        longtable = TRUE,
        linesep   = "",
        # Center the three short columns: Water Class(es), Criterion Value,
        # Criterion Type. align applies to both header and body cells.
        align     = c("l", "l", "l", "c", "c", "l", "c", "l")) %>%
    kable_styling(latex_options = c("striped", "repeat_header")) %>%
    # Fixed widths force wrapping at full font size. LaTeX adds ~0.42cm of
    # padding (2\\tabcolsep) PER column on top of these p{} widths, so eight
    # columns carry ~3.4cm of padding; the content widths below sum to 11.9cm,
    # for ~15.3cm total inside the 16.5cm text block. Parameter and
    # Basis/Notes are the narrow-and-wrapping columns per request.
    column_spec(1, width = "2.4cm") %>%   # Parameter (wraps)
    column_spec(2, width = "1.0cm") %>%   # Code
    column_spec(3, width = "1.6cm") %>%   # Flag Reason (wraps)
    column_spec(4, width = "1.7cm") %>%   # Water Class(es) (wraps, centered)
    column_spec(5, width = "1.5cm") %>%   # Criterion Value (centered)
    column_spec(6, width = "0.8cm") %>%   # Unit
    column_spec(7, width = "1.3cm") %>%   # Criterion Type (centered)
    column_spec(8, width = "1.9cm")       # Basis / Notes (wraps)
  )
} else {
  cat("No manual WQS criteria were entered for this analysis.")
}
```
