---
title: "NPDES Reasonable Potential (RP) Report"
subtitle: "`r paste('Report Generated:', format(Sys.Date(), '%B %d, %Y'))`"
format:
  pdf:
    toc: true
    toc-depth: 1
    number-sections: true
    geometry: margin=1in
    documentclass: article
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{float}
execute:
  echo: false
params:
  permit_id: ""
  facility_name: ""
  date_start: ""
  date_end: ""
  forms: ""
  dilution_ratio: 1
  confidence_level: 0.95
  target_percentile: 0.95
  hardness_mode: "range"
  hardness_value: NA
  rp_path: ""
  coverage_path: ""
  dmr_path: ""
  ph_dmr_path: ""
  temp_dmr_path: ""
  wqs_info_path: ""
  overrides_path: ""
  quick_run: false
  direct_dmr_path: ""    # observations for all rp_method == "direct" parameters
  quick_run_params: ""   # comma-separated list of auto-excluded parameter descs
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = "center", fig.show = "asis",
  # fig.width matches the 6.5in usable text width (letter, 1in margins) so
  # plots render at the correct aspect ratio and never exceed \linewidth.
  # out.width caps the displayed width; fig.height keeps plots compact.
  fig.width = 6.5, fig.height = 3.6, out.width = "\\linewidth"
)

library(readr)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggtext)
library(lubridate)
library(ggforce)
library(ggnewscale)
library(tidyr)

# Inline copy of format_water_class() — keeps the report self-contained when
# rendered from tmp_dir (functions.R is not copied there). If the helper in
# functions.R is updated, mirror the change here.
format_water_class <- function(class, sub_class = NA, sep = " — ") {
  has_sub <- !is.na(sub_class) & nzchar(as.character(sub_class))
  ifelse(has_sub,
         paste0(as.character(class), sep, as.character(sub_class)),
         as.character(class))
}

# Inline copy of format_outfall_label() — same self-containment reason as
# above. Produces "Internal Outfall NNN" or "External Outfall NNN" so the
# type is always visible in the report. Mirror functions.R if this changes.
format_outfall_label <- function(perm_feature_nmbr, perm_feature_type_code = NA_character_) {
  is_internal <- !is.na(perm_feature_type_code) & perm_feature_type_code == "INO"
  ifelse(
    is_internal,
    paste0("Internal Outfall ", perm_feature_nmbr),
    paste0("External Outfall ", perm_feature_nmbr)
  )
}

# LaTeX version: wraps Internal Outfall labels in \textcolor{red}{...} so
# they appear in red in PDF section headings and plot titles.
# External Outfall labels are returned as-is (no colour command needed).
format_outfall_label_latex <- function(perm_feature_nmbr, perm_feature_type_code = NA_character_) {
  lbl <- format_outfall_label(perm_feature_nmbr, perm_feature_type_code)
  is_internal <- !is.na(perm_feature_type_code) & perm_feature_type_code == "INO"
  ifelse(
    is_internal,
    paste0("\\textcolor{red}{", lbl, "}"),
    lbl
  )
}

# Markdown/HTML version: wraps Internal Outfall labels in a red <span> for
# use with ggtext::element_markdown() in ggplot title themes.
format_outfall_label_md <- function(perm_feature_nmbr, perm_feature_type_code = NA_character_) {
  lbl <- format_outfall_label(perm_feature_nmbr, perm_feature_type_code)
  is_internal <- !is.na(perm_feature_type_code) & perm_feature_type_code == "INO"
  ifelse(
    is_internal,
    paste0("<span style='color:red'>", lbl, "</span>"),
    lbl
  )
}

rp_df       <- readr::read_csv(params$rp_path, show_col_types = FALSE) %>%
  dplyr::mutate(CRITERION_ID = as.character(CRITERION_ID))

# Backfill the legacy Method column if the upstream app no longer supplies it.
# The pH section keys its Min/Max criterion lines on Method; rp_operator is
# the authoritative replacement ("<" = floor = Min bound, ">" = ceiling =
# Max bound). Deriving it here keeps the report renderable against either
# vintage of rp_concentration.csv.
if (!"Method" %in% names(rp_df)) {
  rp_df <- rp_df %>%
    dplyr::mutate(
      Method = if ("rp_operator" %in% names(.)) {
        dplyr::case_when(
          is.na(rp_operator) ~ "Concentration",
          rp_operator == "<" ~ "Min",
          rp_operator == ">" ~ "Max",
          TRUE ~ "Concentration"
        )
      } else {
        "Concentration"
      }
    )
}
coverage_df <- if (nzchar(params$coverage_path)) readr::read_csv(params$coverage_path, show_col_types = FALSE) else NULL
dmr_df      <- if (nzchar(params$dmr_path))      readr::read_csv(params$dmr_path,      show_col_types = FALSE) else NULL
# ── crosswalk_v2 helpers (inline; mirror functions.R if changed) ────────────
# Binary criteria store CRITERION_VALUE = 0 so the >-comparison works. They
# are operationalised narrative standards (PRWQS 1303.1.H oil & grease,
# Class SG surfactants), and print as 0 — the narrative basis is carried in
# rp_note rather than replacing the number.
fmt_criterion <- function(value, criterion_form = NULL) {
  ifelse(is.na(value), "--", as.character(value))
}

# "Calculated" = TSD multiplying-factor projection. "Observed" = the value as
# reported, with no MF and no dilution ratio applied. Distinguishing these in
# the report matters: an observed DO minimum must not read as dilution-adjusted.
basis_label <- function(rp_method, rp_operator = NULL) {
  m <- ifelse(is.na(rp_method), "projected", rp_method)
  op <- if (is.null(rp_operator)) rep(NA_character_, length(m)) else rp_operator
  ifelse(
    m == "direct",
    ifelse(!is.na(op) & op == "<", "Observed min", "Observed max"),
    "Calculated"
  )
}

# wqs_info is now the full crosswalk_effective — all columns preserved.
# Each section that uses it selects the columns it needs locally.
wqs_info_df <- if (nzchar(params$wqs_info_path)) {
  readr::read_csv(params$wqs_info_path, show_col_types = FALSE) %>%
    dplyr::mutate(CRITERION_ID = as.character(CRITERION_ID))
} else NULL

# Slim annotation table for the RP breakdown (criterion type columns only)
# Used by the per-parameter detail tables — same purpose as the old wqs_info_df
wqs_annotation_df <- if (!is.null(wqs_info_df)) {
  wqs_info_df %>%
    dplyr::select(
      CRITERION_ID,
      dplyr::any_of(c("CRITERIATYPEAQUAHUMHLTH", "CRITERIATYPEFRESHSALTWATER"))
    ) %>%
    dplyr::distinct()
} else NULL

overrides_df <- if (nzchar(params$overrides_path)) {
  readr::read_csv(params$overrides_path, show_col_types = FALSE)
} else NULL

quick_run       <- isTRUE(params$quick_run)
quick_run_params_vec <- if (nzchar(params$quick_run_params %||% "")) {
  trimws(strsplit(params$quick_run_params, ",")[[1]])
} else character(0)

# Count outfalls (unique perm_feature_nmbr)
outfall_count <- if (!is.null(dmr_df) && "perm_feature_nmbr" %in% names(dmr_df)) {
  dplyr::n_distinct(dmr_df$perm_feature_nmbr)
} else {
  0L
}

# Permit limit history — guarded for standalone runs where ECHO limit columns may be absent
# perm_feature_nmbr is included so the table can be broken down by outfall.
limits_required_cols <- c("parameter_code", "limit_value_nmbr", "limit_unit_desc",
                           "limit_begin_date", "limit_end_date")
# Mass-load unit patterns — these are effluent load (mass/time) measurements that
# cannot be compared to concentration-based WQS criteria and are reported separately.
mass_load_units <- c("kg/d", "kg/day", "lbs/d", "lb/d", "lbs/day", "lb/day",
                     "mg/d", "mg/day", "g/d", "g/day", "ton/d", "ton/day",
                     "kg/mo", "lbs/mo", "kg/yr", "lbs/yr")

limits <- if (!is.null(dmr_df) && all(limits_required_cols %in% names(dmr_df))) {
  base <- dmr_df %>%
    dplyr::select(
      dplyr::all_of(limits_required_cols),
      dplyr::any_of("perm_feature_nmbr")   # include outfall if present
    ) %>%
    tidyr::drop_na(dplyr::all_of(limits_required_cols)) %>%
    dplyr::mutate(
      limit_begin_date = suppressWarnings(dplyr::coalesce(lubridate::ymd(limit_begin_date), lubridate::mdy(limit_begin_date))),
      limit_end_date   = suppressWarnings(dplyr::coalesce(lubridate::ymd(limit_end_date),   lubridate::mdy(limit_end_date))),
      is_mass_load     = tolower(trimws(limit_unit_desc)) %in%
                           tolower(mass_load_units)
    ) %>%
    dplyr::distinct()
  base
} else {
  NULL
}

# Metal criterion IDs and constants (for on-the-fly hardness curves in the report)
metals_ids <- c("79228","79238","79236","79253","79248","79251","79264")

metal_constants <- data.frame(
  CRITERION_ID = metals_ids,
  LC = c(0.7977, 0.8190, 0.8545, 1.273, 0.8460, 1.72, 0.8473),
  RC = c(3.909, 0.6848, 1.702, 4.705, 0.0584, 6.59, 0.884)
)

# Build curve of limits for a given CRITERION_ID across hardness 25..400 (step = 1)
metal_limit_curve <- function(crit_id, h_min = 25, h_max = 400) {
  row <- metal_constants[metal_constants$CRITERION_ID == crit_id, ]
  if (nrow(row) == 0) return(NULL)
  hardness <- seq(h_min, h_max, 1)
  limit <- exp(row$LC * log(hardness) - row$RC)
  dplyr::tibble(CRITERION_ID = crit_id, hardness = hardness, limit = limit)
}

# Safe readers for optional inputs
safe_read_csv <- function(path) {
  if (is.character(path) && nzchar(path) && file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    NULL
  }
}

ph_dmr_df     <- safe_read_csv(params$ph_dmr_path)
temp_dmr_df   <- safe_read_csv(params$temp_dmr_path)
direct_dmr_df <- safe_read_csv(params$direct_dmr_path)

# Normalize dates if present
if (is.data.frame(ph_dmr_df) && "monitoring_period_end_date" %in% names(ph_dmr_df)) {
  if (!inherits(ph_dmr_df$monitoring_period_end_date, "Date")) {
    ph_dmr_df$monitoring_period_end_date <- suppressWarnings(dplyr::coalesce(
      lubridate::ymd(ph_dmr_df$monitoring_period_end_date),
      lubridate::mdy(ph_dmr_df$monitoring_period_end_date)
    ))
  }
}
if (is.data.frame(direct_dmr_df) && "monitoring_period_end_date" %in% names(direct_dmr_df)) {
  if (!inherits(direct_dmr_df$monitoring_period_end_date, "Date")) {
    direct_dmr_df$monitoring_period_end_date <- suppressWarnings(dplyr::coalesce(
      lubridate::ymd(direct_dmr_df$monitoring_period_end_date),
      lubridate::mdy(direct_dmr_df$monitoring_period_end_date)
    ))
  }
}
if (is.data.frame(temp_dmr_df) && "monitoring_period_end_date" %in% names(temp_dmr_df)) {
  if (!inherits(temp_dmr_df$monitoring_period_end_date, "Date")) {
    temp_dmr_df$monitoring_period_end_date <- suppressWarnings(dplyr::coalesce(
      lubridate::ymd(temp_dmr_df$monitoring_period_end_date),
      lubridate::mdy(temp_dmr_df$monitoring_period_end_date)
    ))
  }
}

# ── Shared visual palette ────────────────────────────────────────────────────
# Single source of truth for water class colors and linetypes.
# Used identically in pH, temperature, and concentration-based plots so the
# legend is immediately recognisable across all sections of the report.
#
# Drawing order rule: WQS lines are drawn thickest → thinnest so no line is
# buried under a wider one.  Within WQS lines, Min bounds (lwr_lwd) are drawn
# before Max bounds (upr_lwd); the RWC line is drawn last (on top).
#
wqs_colors <- c(
  "class SB waters"                 = "#1f78b4",  # blue
  "class SD waters"                 = "#ff7f00",  # orange
  "class SD waters- drinking water" = "#33a02c",  # green
  "class SG waters"                 = "#6a3d9a",  # purple
  "class SG waters- drinking water" = "#e31a1c",  # red
  "surface waters"                  = "#b15928",  # brown
  # Sub-class variants (PR class SD: stream vs reservoir/lake). Keyed by the
  # same combined label format_water_class() produces in the table cells, so
  # legends across the report stay consistent.
  "class SD waters \u2014 stream"         = "#cc6600",  # darker orange
  "class SD waters \u2014 reservoir/lake" = "#ffaa55"   # lighter orange
)

wqs_linetypes <- c(
  "class SB waters"                 = "dashed",
  "class SD waters"                 = "dashed",
  "class SD waters- drinking water" = "dotdash",
  "class SG waters"                 = "dashed",
  "class SG waters- drinking water" = "dotdash",
  "surface waters"                  = "longdash",
  "class SD waters \u2014 stream"         = "dashed",
  "class SD waters \u2014 reservoir/lake" = "dashed"
)

# Line widths: Max-bound lines are thinner and drawn on top; Min-bound lines
# are thicker and drawn first so Max is always visible above them.
wqs_lwd_max <- 0.7   # upper limit lines (Max pH, or WQS criterion for conc.)
wqs_lwd_min <- 1.3   # lower limit lines (Min pH)
rwc_color   <- "#2c7fb8"
rwc_lwd     <- 1.0   # RWC line — drawn last, distinct colour keeps it visible
limit_color <- "#d61309"  # hardness curve line colour

# Helper: given a vector of class names present in a plot, return only the
# palette entries that are needed (avoids orphan legend keys).
active_wqs_colors    <- function(classes) wqs_colors[intersect(names(wqs_colors), classes)]
active_wqs_linetypes <- function(classes) wqs_linetypes[intersect(names(wqs_linetypes), classes)]

# Preset point colors by reported unit (dmr_unit_desc) — unchanged from before
unit_levels <- c(
  "deg C","mg/L","SU","mL/L","ug/L","MGD","NTU","col unit (pc)",
  "MF/L","kg/d","#/100mL","%","d","gal/d","MPN/100mL","m3/d"
)
unit_colors <- setNames(
  c(
    "#1b9e77", "#d95f02", "#7570b3", "#e7298a",
    "#66a61e", "#e6ab02", "#a6761d", "#1f78b4",
    "#b2df8a", "#fb9a99", "#e31a1c", "#ff7f00",
    "#6a3d9a", "#b15928", "#33a02c", "#a6cee3"
  ),
  unit_levels
)
```

# Summary

Permit: `r params$permit_id`  
Facility: `r params$facility_name`  
Date range: `r params$date_start` to `r params$date_end`  
Selected forms: `r params$forms`  
Dilution ratio: `r params$dilution_ratio`  
Confidence level: `r params$confidence_level`  
Target percentile: `r params$target_percentile`  
Hardness setting: `r if (params$hardness_mode == "set") paste0("Set (", params$hardness_value, " mg/L as CaCO3)") else "Range"`  

```{r}
#| label: quick-run-warning
#| echo: false
#| results: asis
if (isTRUE(quick_run) && length(quick_run_params_vec) > 0) {
  cat("\n\n> **\\textcolor{red}{WARNING — PRELIMINARY REPORT (Quick Run)}**\n>\n")
  cat("> This report was generated using Quick Run. The following parameters had unresolved\n")
  cat("> data issues and were **automatically excluded** from analysis. Results should be\n")
  cat("> considered preliminary. Re-run with all parameters reviewed for a final report.\n>\n")
  cat(paste0("> *Excluded parameters: ", paste(quick_run_params_vec, collapse = "; "), "*\n\n"))
}
```

## Included Pollutants

The following table shows the available data for each pollutant included in this analysis. The *Basis for Inclusion* column records why each pollutant is present: required by a selected NPDES form (`r params$forms`); carried in because prior DMR data exists even though no selected form requires it (*Existing DMR*); or present only because the user supplied the monitoring data (*Manually Added*). A \ddag\ marker flags parameters whose record includes monitoring data manually added by the user on top of the base DMR record. For a complete list of pollutants that had no associated data, refer to @sec-unavailable.

```{r}
#| label: coverage
#| echo: false

if (!is.null(coverage_df) && nrow(coverage_df) > 0) {

  # Determine which parameter codes had manual WQS entries (for dagger marker)
  manual_codes <- if (!is.null(overrides_df) && nrow(overrides_df) > 0) {
    overrides_df %>%
      dplyr::filter(decision == "include") %>%
      dplyr::pull(parameter_code)
  } else character(0)

  # `User Data` (from coverage.csv): "" none / "partial" some rows
  # user-supplied / "all" every row user-supplied. Older coverage files
  # without the column degrade gracefully to "".
  has_user_col <- "User Data" %in% names(coverage_df)

  top <- coverage_df %>%
    dplyr::filter(`# Samples` > 0) %>%
    dplyr::arrange(Pollutant) %>%
    dplyr::distinct() %>%
    dplyr::mutate(
      # as.character() guards the readr type guess: a coverage.csv where
      # every User Data value is empty parses as an all-NA *logical* column,
      # and coalesce(<logical>, "") is a type error that kills the chunk.
      user_flag = if (has_user_col) {
        dplyr::coalesce(as.character(`User Data`), "")
      } else {
        ""
      },
      # Basis for Inclusion:
      #   * form-required            -> "Form 2C" / "Forms 2C & 2F"
      #   * non-form, base DMR data  -> "Existing DMR"
      #   * non-form, all user rows  -> "Manually Added"
      `Basis for Inclusion` = dplyr::case_when(
        Form == "Not in selected forms" & user_flag == "all" ~ "Manually Added",
        Form == "Not in selected forms" ~ "Existing DMR",
        grepl("[,&]", Form) ~ paste("Forms", Form),
        TRUE ~ paste("Form", Form)
      ),
      # Two independent footnote markers:
      #   dagger  — WQS criterion manually entered (unchanged)
      #   ddagger — record includes user-added monitoring rows, except when
      #             the basis already says "Manually Added" (redundant there)
      Flag = paste0(
        dplyr::if_else(`Ref Param` %in% manual_codes, "\u2020", ""),
        dplyr::if_else(
          user_flag != "" & `Basis for Inclusion` != "Manually Added",
          "\u2021",
          ""
        )
      )
    ) %>%
    dplyr::select(
      Pollutant, `Basis for Inclusion`, `Ref Param`, `# Samples`, Flag
    )

  tbl <- kable(top,
      format    = "latex",
      booktabs  = TRUE,
      longtable = TRUE,
      col.names = c("Pollutant", "Basis for Inclusion", "Ref Param", "# Samples", "")) %>%
    kable_styling(latex_options = c("striped", "repeat_header"))

  # Print each footnote only when its marker actually appears in a cell;
  # an unconditional footnote implies flagged rows that don't exist (and in
  # a run with no overrides and no user-added data, no footnote at all).
  has_dagger  <- any(grepl("\u2020", top$Flag))
  has_ddagger <- any(grepl("\u2021", top$Flag))
  fn_texts <- c(
    if (has_dagger) {
      "WQS criterion manually entered by user; see Manual WQS Overrides section."
    } else {
      NULL
    },
    if (has_ddagger) {
      "Includes monitoring data manually added by the user."
    } else {
      NULL
    }
  )
  fn_marks <- c(
    if (has_dagger) "\u2020" else NULL,
    if (has_ddagger) "\u2021" else NULL
  )
  if (length(fn_texts) > 0) {
    tbl <- tbl %>%
      footnote(symbol = fn_texts, symbol_manual = fn_marks, escape = FALSE)
  }
  tbl

} else {
  cat("No coverage summary available.")
}
```

# RP Findings

RP column: **YES** indicates the assessed value exceeds the criterion.
**DEPENDS** indicates an exceedance was not observed but the tool cannot
conclude no reasonable potential — see the note on the pollutant's section.
**NO** indicates no exceedance.

The Basis column distinguishes a *Calculated* receiving water concentration
(TSD multiplying-factor projection, dilution ratio applied) from an
*Observed* maximum or minimum compared directly to the criterion with no
projection and no dilution. Values rounded to 2 decimals.

```{r}
#| label: rp-table
#| echo: false

rp_df %>%
  mutate(Basis = basis_label(
           if ("rp_method" %in% names(.)) rp_method else NA_character_,
           if ("rp_operator" %in% names(.)) rp_operator else NULL
         ),
         RWC_rs = round(RWC_rs, 2),
         CRITERION_VALUE = round(CRITERION_VALUE,2),
         CRITERION_ID = gsub("_", "\\\\_", CRITERION_ID),
         NPDES_Pollutant = gsub("_", "\\\\_", NPDES_Pollutant)) %>%
  drop_na(RWC_rs) %>%
  mutate(USE_CLASS_NAME_LOCATION_ETC = format_water_class(
           USE_CLASS_NAME_LOCATION_ETC,
           if ("sub_class" %in% names(.)) sub_class else NA_character_
         ),
         # Internal Outfalls: red bold text via cell_spec so they stay
         # visibly distinct from External Outfalls in the PDF.
         perm_feature_nmbr = {
           lbl <- format_outfall_label(
             perm_feature_nmbr,
             if ("perm_feature_type_code" %in% names(.)) perm_feature_type_code else NA_character_
           )
           is_int <- !is.na(perm_feature_type_code) & perm_feature_type_code == "INO"
           ifelse(
             is_int,
             kableExtra::cell_spec(lbl, format = "latex", color = "red", bold = TRUE),
             lbl
           )
         }) %>%
  select(CRITERION_ID, NPDES_Pollutant, perm_feature_nmbr, USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE, RWC_rs, Basis, RP) %>%
  # Severity order YES > DEPENDS > NO. A plain desc() on the character column
  # would sort YES, NO, DEPENDS and bury the determinations needing review.
  arrange(desc(factor(RP, levels = c("NO", "DEPENDS", "YES"), ordered = TRUE)),
          NPDES_Pollutant) %>%
  distinct() %>%
  kable(
    col.names = c("WQS ID","Pollutant","Outfall","Water Class","WQS Criterion","Value","Basis","RP"),
    format    = "latex",
    booktabs  = TRUE,
    longtable = TRUE,
    escape    = FALSE
  ) %>%
  kable_styling(latex_options = c("striped", "repeat_header"),
                font_size = 9) %>%
  # Explicit widths so all 7 columns fit within the 6.5in text block.
  # Sum ≈ 13.6cm (5.4in), leaving margin for inter-column padding.
  column_spec(1, width = "1.1cm") %>%   # WQS ID
  column_spec(2, width = "3.2cm") %>%   # Pollutant
  column_spec(3, width = "2.0cm") %>%   # Outfall (wraps for plain + coloured labels)
  column_spec(4, width = "2.6cm") %>%   # Water Class
  column_spec(5, width = "1.5cm") %>%   # WQS Criterion
  column_spec(6, width = "1.3cm") %>%   # Value (calculated RWC or observed)
  column_spec(7, width = "1.7cm") %>%   # Basis
  column_spec(8, width = "1.4cm")       # RP
```

# Non-Concentration Based Measures

```{r}
#| label: receiving-water
#| echo: false
#| results: asis

# Helper for existence checks
has_rows <- function(df) is.data.frame(df) && nrow(df) > 0

if (is.null(rp_df)) {
  cat("No RP results are available; skipping receiving water sections.\n\n")
} else {
  # All direct-comparison parameters other than pH/temperature, which keep
  # their bespoke branches below. These share one generic table+plot shape.
  direct_extra <- if (has_rows(direct_dmr_df)) {
    direct_dmr_df %>%
      dplyr::filter(!parameter_code %in% c("00400", "00010"))
  } else {
    NULL
  }

  # Build outfall list from whatever is available
  outfalls <- unique(c(
    if (has_rows(ph_dmr_df))   ph_dmr_df$perm_feature_nmbr else character(0),
    if (has_rows(temp_dmr_df)) temp_dmr_df$perm_feature_nmbr else character(0),
    if (has_rows(direct_extra)) direct_extra$perm_feature_nmbr else character(0),
    if (has_rows(rp_df))       rp_df$perm_feature_nmbr else character(0)
  ))
  outfalls <- sort(unique(outfalls[!is.na(outfalls)]))

  if (length(outfalls) == 0 ||
      (!has_rows(ph_dmr_df) && !has_rows(temp_dmr_df) && !has_rows(direct_extra))) {
    cat("No non-concentration receiving-water data to report.\n\n")
  } else {

    for (of in outfalls) {
      # PAGE-BREAK SCOPING: each PARAMETER block below (heading + note +
      # tables + plot) is wrapped in its own \begin{samepage} group. A
      # samepage group SMALLER than one page relocates intact to the next
      # page when it doesn't fit — clean breaks between parameters, never
      # inside one. An earlier version wrapped the entire OUTFALL in one
      # group; with pH + turbidity + DO per outfall that group exceeded a
      # full page, samepage forbade the breaks LaTeX needed, and the
      # overfilled page clipped the headings and plots typeset past the
      # page bottom. Keep the groups per-parameter: if a single parameter
      # block ever grows taller than one page, the clipping returns.

      # Compute display labels for this outfall — plain, LaTeX-coloured, and
      # markdown-coloured. Internal Outfalls appear in red in section headings
      # (\textcolor) and ggplot titles (element_markdown span).
      of_type <- if ("perm_feature_type_code" %in% names(rp_df)) {
        rp_df %>%
          dplyr::filter(perm_feature_nmbr == of) %>%
          dplyr::pull(perm_feature_type_code) %>%
          dplyr::first()
      } else {
        NA_character_
      }
      of_label       <- format_outfall_label(of, of_type)
      of_label_latex <- format_outfall_label_latex(of, of_type)
      of_label_md    <- format_outfall_label_md(of, of_type)

      # ── pH ──────────────────────────────────────────────────────────────────
      if (has_rows(ph_dmr_df)) {
        ph_obs <- ph_dmr_df %>%
          dplyr::filter(perm_feature_nmbr == of) %>%
          dplyr::arrange(monitoring_period_end_date)

        if (nrow(ph_obs) > 0) {
          cat("\n\n```{=latex}\n\\begin{samepage}\n```\n\n")
          cat(sprintf("## %s — pH\n\n", of_label_latex))

          # Summary table
          ph_summ <- ph_obs %>%
            dplyr::summarise(
              max_pH = max(dmr_value_nmbr, na.rm = TRUE),
              min_pH = min(dmr_value_nmbr, na.rm = TRUE)
            )
          print(knitr::kable(ph_summ, format = "latex", booktabs = TRUE,
                             col.names = c("Max pH", "Min pH")))
          cat("\n\n")

          # WQS lines from rp_df for pH at this outfall.
          # Each row has USE_CLASS_NAME_LOCATION_ETC + Method (Max / Min).
          # We build a data frame with one row per line, tagged so we can
          # control drawing order: Min lines first (thick), Max lines on top (thin).
          wqs_ph_raw <- rp_df %>%
            dplyr::filter(parameter_code == "00400", perm_feature_nmbr == of) %>%
            dplyr::select(USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE, Method) %>%
            dplyr::distinct() %>%
            dplyr::filter(is.finite(CRITERION_VALUE))

          # Build combined line-key used for colour + linetype scale lookup
          wqs_ph <- wqs_ph_raw %>%
            dplyr::mutate(
              line_key = USE_CLASS_NAME_LOCATION_ETC,
              # linewidth: Min bound drawn thicker first, Max drawn thinner on top
              lwd = dplyr::if_else(Method == "Min", wqs_lwd_min, wqs_lwd_max),
              # draw order: Min rows first so Max is rendered on top
              draw_order = dplyr::if_else(Method == "Min", 1L, 2L)
            ) %>%
            dplyr::arrange(draw_order)

          present_classes <- unique(wqs_ph$line_key)

          # y-axis lower bound: a unit below the smallest value among obs + WQS
          min_value <- min(c(ph_obs$dmr_value_nmbr, wqs_ph$CRITERION_VALUE),
                          na.rm = TRUE) - 1

          p <- ggplot2::ggplot(
            ph_obs,
            ggplot2::aes(x = monitoring_period_end_date, y = dmr_value_nmbr)
          ) +
            ggplot2::geom_line(color = "lightgrey", linetype = "dotted",
                               linewidth = 0.5)

          # Draw WQS lines in draw_order (Min first, Max on top).
          # Each row gets its own linewidth via a loop so we can vary lwd
          # per row — geom_hline does not support a lwd aesthetic directly.
          if (nrow(wqs_ph) > 0) {
            for (i in seq_len(nrow(wqs_ph))) {
              row_i <- wqs_ph[i, ]
              p <- p + ggplot2::geom_hline(
                yintercept = row_i$CRITERION_VALUE,
                color      = wqs_colors[row_i$line_key],
                linetype   = wqs_linetypes[row_i$line_key],
                linewidth  = row_i$lwd,
                inherit.aes = FALSE
              )
            }
          }

          p <- p +
            ggplot2::geom_point(size = 1.8, color = "#7570b3") +
            ggplot2::scale_y_continuous(limits = c(min_value, NA)) +
            ggplot2::labs(
              title    = sprintf("pH over time — %s", of_label_md),
              x        = "Date",
              y        = "pH (SU)"
            ) +
            ggplot2::theme_minimal() +
            ggplot2::theme(
              plot.title      = ggtext::element_markdown(),
              legend.position = "bottom",
              legend.box      = "vertical",
              legend.text     = ggplot2::element_text(size = 7),
              legend.title    = ggplot2::element_text(size = 8)
            )
          # We show Max lines in the legend (representative of each class).
          if (length(present_classes) > 0) {
            legend_df <- wqs_ph %>%
              dplyr::filter(draw_order == 2L) %>%   # Max rows only for legend
              dplyr::distinct(line_key, .keep_all = TRUE)

            p <- p +
              ggplot2::geom_hline(
                data        = legend_df,
                ggplot2::aes(yintercept = CRITERION_VALUE,
                             color      = line_key,
                             linetype   = line_key),
                linewidth   = wqs_lwd_max,
                inherit.aes = FALSE,
                show.legend = TRUE
              ) +
              ggplot2::scale_color_manual(
                name   = "Water Class",
                values = active_wqs_colors(present_classes)
              ) +
              ggplot2::scale_linetype_manual(
                name   = "Water Class",
                values = active_wqs_linetypes(present_classes)
              )
          }

          knitr::opts_current$set(out.width = "\\linewidth",
                                  fig.height = 3.4,
                                  fig.align = "center", fig.show = "asis")
          print(p)
          cat("\n\n")

          # RP table under pH plot
          ph_rp_tbl <- wqs_ph_raw %>%
            dplyr::left_join(
              rp_df %>%
                dplyr::filter(parameter_code == "00400", perm_feature_nmbr == of) %>%
                dplyr::select(USE_CLASS_NAME_LOCATION_ETC, CRITERION_ID, Method, RP) %>%
                dplyr::distinct(),
              by = c("USE_CLASS_NAME_LOCATION_ETC", "Method")
            ) %>%
            dplyr::mutate(
              Observed = dplyr::if_else(
                Method == "Max",
                round(max(ph_obs$dmr_value_nmbr, na.rm = TRUE), 2),
                round(min(ph_obs$dmr_value_nmbr, na.rm = TRUE), 2)
              ),
              CRITERION_VALUE = round(CRITERION_VALUE, 2),
              # Show sub_class in the Water Class cell when present. pH
              # criteria don't currently have sub_class entries in PR, but
              # the column is generic so this handles future cases.
              USE_CLASS_NAME_LOCATION_ETC = format_water_class(
                USE_CLASS_NAME_LOCATION_ETC,
                if ("sub_class" %in% names(.)) sub_class else NA_character_
              )
            ) %>%
            dplyr::select(CRITERION_ID, USE_CLASS_NAME_LOCATION_ETC,
                          Method, CRITERION_VALUE, Observed, RP)

          if (nrow(ph_rp_tbl) > 0) {
            tbl <- knitr::kable(
              ph_rp_tbl,
              format    = "latex",
              booktabs  = TRUE,
              col.names = c("WQS ID", "Water Class", "Bound",
                            "WQS Criterion", "Observed pH", "RP")
            ) %>% kableExtra::kable_styling(latex_options = c("HOLD_position"))
            print(tbl)
            cat("\n\n")
          }
          cat("\n\n```{=latex}\n\\end{samepage}\n```\n\n")
        }
      } # end pH

      # ── Temperature ─────────────────────────────────────────────────────────
      if (has_rows(temp_dmr_df)) {
        t_obs <- temp_dmr_df %>%
          dplyr::filter(perm_feature_nmbr == of) %>%
          dplyr::arrange(monitoring_period_end_date)

        if (nrow(t_obs) > 0) {
          cat("\n\n```{=latex}\n\\begin{samepage}\n```\n\n")
          # Use a sub-heading if pH was also present for this outfall
          if (has_rows(ph_dmr_df) &&
              nrow(dplyr::filter(ph_dmr_df, perm_feature_nmbr == of)) > 0) {
            cat(sprintf("### Temperature — %s\n\n", of_label_latex))
          } else {
            cat(sprintf("## %s — Temperature\n\n", of_label_latex))
          }

          # Summary table
          t_summ <- t_obs %>%
            dplyr::summarise(Max_Temperature_C = max(dmr_value_nmbr, na.rm = TRUE))
          print(knitr::kable(t_summ, format = "latex", booktabs = TRUE,
                             col.names = c("Max Temperature (°C)")))
          cat("\n\n")

          # WQS lines for temperature — temperature is a single upper bound per
          # class (no Min/Max pair), so no linewidth split needed.
          # Still ordered so thicker lines are drawn first.
          wqs_t <- rp_df %>%
            dplyr::filter(parameter_code == "00010", perm_feature_nmbr == of) %>%
            dplyr::select(USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE) %>%
            dplyr::distinct() %>%
            dplyr::filter(is.finite(CRITERION_VALUE)) %>%
            # Draw thicker lines first (more classes = draw in consistent order)
            dplyr::arrange(USE_CLASS_NAME_LOCATION_ETC)

          present_t_classes <- unique(wqs_t$USE_CLASS_NAME_LOCATION_ETC)

          p2 <- ggplot2::ggplot(
            t_obs,
            ggplot2::aes(x = monitoring_period_end_date, y = dmr_value_nmbr)
          ) +
            ggplot2::geom_line(color = "lightgrey", linetype = "dotted",
                               linewidth = 0.5)

          # Draw WQS lines individually to use consistent lwd
          if (nrow(wqs_t) > 0) {
            for (i in seq_len(nrow(wqs_t))) {
              row_i <- wqs_t[i, ]
              p2 <- p2 + ggplot2::geom_hline(
                yintercept  = row_i$CRITERION_VALUE,
                color       = wqs_colors[row_i$USE_CLASS_NAME_LOCATION_ETC],
                linetype    = wqs_linetypes[row_i$USE_CLASS_NAME_LOCATION_ETC],
                linewidth   = wqs_lwd_max,
                inherit.aes = FALSE
              )
            }
          }

          p2 <- p2 +
            ggplot2::geom_point(size = 1.8, color = "#1b9e77") +
            ggplot2::scale_y_continuous(limits = c(0, NA)) +
            ggplot2::labs(
              title = sprintf("Temperature over time — %s", of_label_md),
              x     = "Date",
              y     = "Temperature (°C)"
            ) +
            ggplot2::theme_minimal() +
            ggplot2::theme(
              plot.title      = ggtext::element_markdown(),
              legend.position = "bottom",
              legend.box      = "vertical",
              legend.text     = ggplot2::element_text(size = 7),
              legend.title    = ggplot2::element_text(size = 8)
            )

          # Legend via zero-alpha ghost layer
          if (length(present_t_classes) > 0) {
            p2 <- p2 +
              ggplot2::geom_hline(
                data        = wqs_t,
                ggplot2::aes(yintercept  = CRITERION_VALUE,
                             color       = USE_CLASS_NAME_LOCATION_ETC,
                             linetype    = USE_CLASS_NAME_LOCATION_ETC),
                linewidth   = wqs_lwd_max,
                inherit.aes = FALSE,
                show.legend = TRUE
              ) +
              ggplot2::scale_color_manual(
                name   = "Water Class",
                values = active_wqs_colors(present_t_classes)
              ) +
              ggplot2::scale_linetype_manual(
                name   = "Water Class",
                values = active_wqs_linetypes(present_t_classes)
              )
          }

          knitr::opts_current$set(out.width = "\\linewidth",
                                  fig.height = 3.4,
                                  fig.align = "center", fig.show = "asis")
          print(p2)
          cat("\n\n")

          # Temperature RP table
          temp_rp_tbl <- rp_df %>%
            dplyr::filter(parameter_code == "00010", perm_feature_nmbr == of) %>%
            dplyr::select(CRITERION_ID, USE_CLASS_NAME_LOCATION_ETC,
                          CRITERION_VALUE, RP,
                          dplyr::any_of("sub_class")) %>%
            dplyr::distinct() %>%
            dplyr::mutate(
              Observed_Max = round(max(t_obs$dmr_value_nmbr, na.rm = TRUE), 2),
              CRITERION_VALUE = round(CRITERION_VALUE, 2),
              # Same combined-display pattern as the concentration and pH
              # tables; future-proofs the temperature table for any
              # jurisdiction that introduces a sub_class for temperature.
              USE_CLASS_NAME_LOCATION_ETC = format_water_class(
                USE_CLASS_NAME_LOCATION_ETC,
                if ("sub_class" %in% names(.)) sub_class else NA_character_
              )
            ) %>%
            dplyr::select(-dplyr::any_of("sub_class"))

          if (nrow(temp_rp_tbl) > 0) {
            tbl <- knitr::kable(
              temp_rp_tbl,
              format    = "latex",
              booktabs  = TRUE,
              col.names = c("WQS ID", "Water Class", "Limit (°C)",
                            "RP", "Observed Max (°C)")
            ) %>% kableExtra::kable_styling(latex_options = c("HOLD_position"))
            print(tbl)
            cat("\n\n")
          }
          cat("\n\n```{=latex}\n\\end{samepage}\n```\n\n")
        }
      } # end temperature

      # ── Other direct-comparison parameters ──────────────────────────────────
      # DO, color, turbidity, Enterococci, oil & grease. One table + one plot
      # each, driven by rp_operator (">" ceiling / "<" floor) and the same
      # log-axis + unit-sanity flag used in the app.
      if (has_rows(direct_extra)) {
        de_of <- direct_extra %>%
          dplyr::filter(perm_feature_nmbr == of)

        for (pcode in sort(unique(de_of$parameter_code))) {
          d_obs <- de_of %>%
            dplyr::filter(parameter_code == pcode) %>%
            dplyr::arrange(monitoring_period_end_date)
          if (nrow(d_obs) == 0) next

          d_rows <- rp_df %>%
            dplyr::filter(parameter_code == pcode, perm_feature_nmbr == of)
          if (nrow(d_rows) == 0) next

          op_dir <- if ("rp_operator" %in% names(d_rows)) {
            d_rows$rp_operator[1]
          } else {
            ">"
          }
          is_floor <- !is.na(op_dir) && op_dir == "<"
          d_name <- if ("criterion_label" %in% names(d_rows) &&
                        !is.na(d_rows$criterion_label[1]) &&
                        nzchar(d_rows$criterion_label[1])) {
            d_rows$criterion_label[1]
          } else {
            d_rows$NPDES_Pollutant[1]
          }
          d_name <- gsub("_", "\\\\_", d_name)

          cat("\n\n```{=latex}\n\\begin{samepage}\n```\n\n")
          cat(sprintf("### %s — %s\n\n", d_name, of_label_latex))

          # Criterion note, if any (Streeter-Phelps, Enterococci interval, etc.)
          if ("rp_note" %in% names(d_rows)) {
            nt <- unique(d_rows$rp_note)
            nt <- nt[!is.na(nt) & nzchar(nt)]
            if (length(nt) > 0) {
              cat(sprintf("> **\\textcolor{orange}{Note}** — %s\n\n",
                          gsub("_", "\\\\_", nt[1])))
            }
          }

          # Observed summary: min for floors, max for ceilings.
          d_val <- if (is_floor) {
            min(d_obs$dmr_value_nmbr, na.rm = TRUE)
          } else {
            max(d_obs$dmr_value_nmbr, na.rm = TRUE)
          }

          d_tbl <- d_rows %>%
            dplyr::select(CRITERION_ID, USE_CLASS_NAME_LOCATION_ETC,
                          CRITERION_VALUE, RP, dplyr::any_of("sub_class")) %>%
            dplyr::distinct() %>%
            dplyr::mutate(
              Observed = round(d_val, 4),
              CRITERION_VALUE = round(CRITERION_VALUE, 4),
              USE_CLASS_NAME_LOCATION_ETC = format_water_class(
                USE_CLASS_NAME_LOCATION_ETC,
                if ("sub_class" %in% names(.)) sub_class else NA_character_
              )
            ) %>%
            dplyr::select(-dplyr::any_of("sub_class"))

          print(knitr::kable(
            d_tbl, format = "latex", booktabs = TRUE,
            col.names = c("WQS ID", "Water Class", "WQS Criterion", "RP",
                          if (is_floor) "Observed Min" else "Observed Max")
          ) %>% kableExtra::kable_styling(latex_options = c("HOLD_position")))
          cat("\n\n")

          # Plot: observations + criterion line(s), log axis + unit flag.
          wqs_d <- d_rows %>%
            dplyr::select(USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE) %>%
            dplyr::distinct() %>%
            dplyr::filter(is.finite(CRITERION_VALUE))

          pos_d  <- d_obs$dmr_value_nmbr[is.finite(d_obs$dmr_value_nmbr) &
            d_obs$dmr_value_nmbr > 0]
          crit_d <- suppressWarnings(min(
            wqs_d$CRITERION_VALUE[is.finite(wqs_d$CRITERION_VALUE) &
              wqs_d$CRITERION_VALUE > 0]
          ))
          oom_d <- if (length(pos_d) > 0 && is.finite(crit_d) && crit_d > 0) {
            max(pos_d) / crit_d
          } else {
            NA_real_
          }
          log_d <- is.finite(oom_d) && oom_d >= 100

          # Non-detects arrive as NODI B/Q rewritten to 0 by the DMR loader;
          # literal reported zeros land in the same place. log10 drops both
          # silently, hiding results the permit writer needs. Plot them at a
          # floor -- the smallest positive value in the series, the closest
          # available proxy for the reporting limit -- as open down-triangles
          # so they cannot be read as measurements. The caption discloses it.
          d_obs$is_nd_d <- is.finite(d_obs$dmr_value_nmbr) &
            d_obs$dmr_value_nmbr <= 0
          nd_floor_d <- if (length(pos_d) > 0) min(pos_d) else NA_real_
          nd_at_floor_d <- log_d && any(d_obs$is_nd_d) && is.finite(nd_floor_d)
          d_obs$y_plot_d <- ifelse(
            d_obs$is_nd_d & nd_at_floor_d,
            nd_floor_d,
            d_obs$dmr_value_nmbr
          )
          nd_cap_d <- if (nd_at_floor_d) {
            paste0("Open triangles are non-detects, plotted at ",
                   signif(nd_floor_d, 3), ".")
          } else if (any(d_obs$is_nd_d)) {
            "Open triangles are non-detects."
          } else {
            NA_character_
          }

          p_d <- ggplot2::ggplot(
            d_obs,
            ggplot2::aes(x = monitoring_period_end_date, y = y_plot_d)
          ) +
            ggplot2::geom_line(color = "lightgrey", linetype = "dotted",
                               linewidth = 0.5) +
            ggplot2::geom_point(
              data = d_obs[!d_obs$is_nd_d, , drop = FALSE],
              color = "#1b9e77", alpha = 0.85
            )

          if (any(d_obs$is_nd_d)) {
            p_d <- p_d + ggplot2::geom_point(
              data  = d_obs[d_obs$is_nd_d, , drop = FALSE],
              shape = 6, color = "grey30", size = 2.2, stroke = 0.9
            )
          }

          if (nrow(wqs_d) > 0) {
            p_d <- p_d + ggplot2::geom_hline(
              data = wqs_d,
              ggplot2::aes(yintercept = CRITERION_VALUE,
                           color = USE_CLASS_NAME_LOCATION_ETC),
              linetype = "dashed", linewidth = 0.7, inherit.aes = FALSE
            ) +
              ggplot2::scale_color_manual(
                name = "Water Class",
                values = active_wqs_colors(
                  unique(wqs_d$USE_CLASS_NAME_LOCATION_ETC)
                )
              )
          }

          p_d <- p_d +
            {
              if (log_d) ggplot2::scale_y_log10()
              else ggplot2::scale_y_continuous()
            } +
            ggplot2::labs(
              x = "Monitoring date",
              y = paste0(d_rows$UNIT_NAME[1], if (log_d) " (log scale)" else ""),
              caption = if (!is.na(nd_cap_d)) nd_cap_d else NULL
            ) +
            ggplot2::theme_minimal() +
            ggplot2::theme(
              legend.position = "bottom",
              plot.caption = ggplot2::element_text(hjust = 0, color = "grey30",
                                                   size = 8)
            )

          print(p_d)
          cat("\n\n")
          cat("\n\n```{=latex}\n\\end{samepage}\n```\n\n")
        }
      } # end other direct params

    } # end outfall loop
  }

} # end rp_df is not NULL
```




# Concentration Based Measures

## RP Equations

The following equations estimate the Receiving Water Concentration (RWC) for
parameters assessed by projection. They do **not** apply to parameters
assessed by direct comparison — pH, temperature, dissolved oxygen,
Enterococci, color, turbidity and oil & grease — whose sections report an
observed maximum or minimum instead, with no multiplying factor and no
dilution ratio. Each such section states this explicitly.

$$\sigma_g = \sqrt{\ln(\mathrm{CV}^2 + 1)}$$
$$p = (1 - \mathrm{CL})^{1/n}$$
$$z_p = \Phi^{-1}(p),\; z_{\text{target}} = \Phi^{-1}(\mathrm{target})$$
$$\mathrm{MF} = \exp\big((z_{\text{target}} - z_p)\,\sigma_g\big)$$

$$\text{RWC} = \frac{\max(x) \times MF}{DR}$$

```{r}
#| label: per-pollutant
#| echo: false
#| results: asis

if (is.null(dmr_df) || nrow(dmr_df) == 0) {
  cat("No DMR data were provided; per-pollutant sections are omitted.")
} else {
  # Direct-comparison parameters have no projected RWC but must still get a
  # section — this is why color, turbidity, DO and Enterococci previously
  # appeared in the data but never in the report.
  pollutants <- rp_df %>%
    dplyr::filter(
      (!is.na(n_used) & n_used > 0) |
        is.finite(RWC_rs) |
        (!is.na(rp_method) & rp_method == "direct")
    ) %>%
    dplyr::distinct(NPDES_Pollutant) %>%
    dplyr::pull()

  for (poll in pollutants) {
    sub_all <- rp_df %>% dplyr::filter(NPDES_Pollutant == poll)
    param_codes <- sub_all %>% dplyr::distinct(parameter_code) %>% dplyr::pull()
    param_code <- if (length(param_codes) > 0) param_codes[1] else NA_character_

    cat(sprintf("\n\n## %s\n\n", poll))

    # ── Criterion note (rp_note) ──────────────────────────────────────────
    # Carries criterion-specific guidance from the crosswalk: the DO
    # Streeter-Phelps data gap, the Enterococci 90-day interval caveat, the
    # in-stream basis of the temperature criterion, the narrative basis of
    # the oil & grease detection test.
    if ("rp_note" %in% names(sub_all)) {
      note_txt <- sub_all %>%
        dplyr::pull(rp_note) %>%
        unique()
      note_txt <- note_txt[!is.na(note_txt) & nzchar(note_txt)]
      if (length(note_txt) > 0) {
        cat(sprintf(
          "> **\\textcolor{orange}{Note}** — %s\n\n",
          gsub("_", "\\\\_", note_txt[1])
        ))
      }
    }

    # ── Reason Included line ───────────────────────────────────────────────
    # Mirrors the Basis for Inclusion mapping in the Included Pollutants
    # table: form-required, Existing DMR, or Manually Added — with a
    # supplemental note when user-added rows sit on top of a base record.
    if (!is.null(coverage_df) && "Form" %in% names(coverage_df)) {
      cov_row <- coverage_df %>%
        dplyr::filter(Pollutant == poll) %>%
        dplyr::slice(1)
      form_val <- if (nrow(cov_row) > 0) cov_row$Form else NA_character_
      user_val <- if (nrow(cov_row) > 0 && "User Data" %in% names(cov_row)) {
        dplyr::coalesce(cov_row$`User Data`, "")
      } else {
        ""
      }
      if (!is.null(form_val) && !is.na(form_val) && nchar(form_val) > 0) {
        basis_is_manual <- form_val == "Not in selected forms" &&
          user_val == "all"
        if (basis_is_manual) {
          cat("*Reason Included: Monitoring data manually added by the user.*\n\n")
        } else if (form_val == "Not in selected forms") {
          cat("*Reason Included: Not associated with selected forms — prior DMR data exists.*\n\n")
        } else {
          cat(sprintf("*Reason Included: Form(s) %s selected.*\n\n", form_val))
        }
        if (user_val != "" && !basis_is_manual) {
          cat("*Includes monitoring data manually added by the user.*\n\n")
        }
      }
    }
    # ── Manual WQS callout ────────────────────────────────────────────────
    if (!is.null(overrides_df) && !is.na(param_code)) {
      ov_row <- overrides_df %>%
        dplyr::filter(parameter_code == param_code, decision == "include") %>%
        dplyr::slice(1)
      if (nrow(ov_row) > 0) {
        cat("> **\\textcolor{orange}{Manual WQS Applied}** — ",
            "Criterion value manually entered by user. ",
            "See *Manual WQS Overrides* section for full details.\n\n",
            sep = "")
      }
    }

    outfalls_for_poll <- if ("perm_feature_nmbr" %in% names(sub_all)) {
      unique(sub_all$perm_feature_nmbr)
    } else if (!is.null(dmr_df) && "perm_feature_nmbr" %in% names(dmr_df)) {
      unique(dplyr::filter(dmr_df, parameter_code == param_code)$perm_feature_nmbr)
    } else NA_character_
    outfalls_for_poll <- outfalls_for_poll[!is.na(outfalls_for_poll)]
    if (length(outfalls_for_poll) == 0) outfalls_for_poll <- NA_character_

    show_outfall_subheads <- is.numeric(outfall_count) && outfall_count > 1

    for (of in outfalls_for_poll) {
      # \nopagebreak[4] strongly discourages a page break immediately before
      # the outfall subheading, keeping it with the first line of content.
      # Unlike \needspace, this requires no extra LaTeX package.
      cat("\n\n\\nopagebreak[4]\n\n")

      # Display labels — plain, LaTeX-coloured, and markdown-coloured variants.
      of_type <- if ("perm_feature_type_code" %in% names(rp_df) && !is.na(of)) {
        rp_df %>%
          dplyr::filter(perm_feature_nmbr == of) %>%
          dplyr::pull(perm_feature_type_code) %>%
          dplyr::first()
      } else {
        NA_character_
      }
      of_label       <- format_outfall_label(of, of_type)
      of_label_latex <- format_outfall_label_latex(of, of_type)
      of_label_md    <- format_outfall_label_md(of, of_type)

      if (show_outfall_subheads && !is.na(of)) cat(sprintf("### %s\n\n", of_label_latex))

      # ── Permit Limit History (per outfall) ────────────────────────────────
      # Limits can vary by outfall, so the table is scoped to the current outfall.
      # Mass-load limits (e.g. kg/d) cannot be compared to concentration-based
      # WQS criteria and are listed separately.
      if (!is.null(limits) && is.data.frame(limits) && !is.na(param_code)) {
        lim_base <- limits %>%
          dplyr::filter(parameter_code == param_code)
        # Filter to this outfall when the column is present
        if ("perm_feature_nmbr" %in% names(lim_base) && !is.na(of)) {
          lim_base <- lim_base %>% dplyr::filter(perm_feature_nmbr == of)
        }

        # Concentration-based limits
        lim_conc <- lim_base %>%
          dplyr::filter(!is_mass_load) %>%
          dplyr::transmute(
            `Effluent Limit` = limit_value_nmbr,
            `Unit`           = limit_unit_desc,
            `Limit Start`    = format(limit_begin_date, "%m/%d/%Y"),
            `Limit End`      = format(limit_end_date,   "%m/%d/%Y")
          ) %>%
          dplyr::distinct()

        if (nrow(lim_conc) > 0) {
          cat("**Permit Effluent Limit History**\n\n")
          kable(lim_conc,
                format    = "latex",
                booktabs  = TRUE,
                col.names = c("Effluent Limit", "Unit", "Limit Start", "Limit End")) %>%
            kable_styling(latex_options = c("HOLD_position")) %>%
            print()
          cat("\n\n")
        }

        # Mass-load limits (reported separately — not used in WQS comparison)
        lim_mass <- lim_base %>%
          dplyr::filter(is_mass_load) %>%
          dplyr::transmute(
            `Effluent Limit` = limit_value_nmbr,
            `Unit`           = limit_unit_desc,
            `Limit Start`    = format(limit_begin_date, "%m/%d/%Y"),
            `Limit End`      = format(limit_end_date,   "%m/%d/%Y")
          ) %>%
          dplyr::distinct()

        if (nrow(lim_mass) > 0) {
          cat("**Mass-Load Effluent Limits (not used in WQS analysis)**\n\n")
          cat(paste0("The following limits are expressed as mass-load quantities ",
                     "(mass per unit time) and cannot be directly compared to the ",
                     "concentration-based WQS criteria used in this analysis. ",
                     "They are provided for reference only.\n\n"))
          kable(lim_mass,
                format    = "latex",
                booktabs  = TRUE,
                col.names = c("Effluent Limit", "Unit", "Limit Start", "Limit End")) %>%
            kable_styling(latex_options = c("HOLD_position")) %>%
            print()
          cat("\n\n")
        }
      }

      sub_rp <- if (!is.na(of)) sub_all %>% dplyr::filter(perm_feature_nmbr == of) else sub_all

      obs <- dmr_df %>%
        dplyr::filter(parameter_code == param_code) %>%
        { if (!is.na(of) && "perm_feature_nmbr" %in% names(.))
            dplyr::filter(., perm_feature_nmbr == of) else . } %>%
        dplyr::mutate(
          # Handle both ISO YYYY-MM-DD (SQLite) and MM/DD/YYYY (ECHO legacy)
          monitoring_period_end_date = suppressWarnings(dplyr::coalesce(
            lubridate::ymd(monitoring_period_end_date),
            lubridate::mdy(monitoring_period_end_date)
          ))
        ) %>%
        dplyr::filter(!is.na(monitoring_period_end_date)) %>%
        dplyr::arrange(monitoring_period_end_date)

      n_samples <- nrow(obs)
      unit_conv_flag <- if ("Conv_Flag" %in% names(obs))
        any(obs$Conv_Flag == "PASS", na.rm = TRUE) else NA
      cat(sprintf("- Samples: %d\n\n", n_samples))
      if (!is.na(unit_conv_flag))
        cat(sprintf("- Units converted: %s\n\n",
                    if (unit_conv_flag) "Yes" else "No"))

      # RWC and WQS lines for this outfall
      rwc_val <- sub_rp %>%
        dplyr::distinct(RWC_rs) %>% dplyr::pull(RWC_rs) %>% .[is.finite(.)]
      rwc_val <- if (length(rwc_val) > 0) rwc_val[1] else NA_real_

      # wqs_lines drives the dashed criterion lines on the per-pollutant
      # plot AND keys their legend entries via wqs_colors / wqs_linetypes.
      # When a parameter has multiple criteria in the same base class that
      # differ by sub_class (e.g., phosphorus SD streams 160 ug/L vs SD
      # reservoir/lake 26 ug/L), we substitute the combined display label
      # so each criterion gets its own legend entry and its own palette
      # slot. Rows without sub_class fall through with the base class.
      wqs_lines <- sub_rp %>%
        dplyr::mutate(
          USE_CLASS_NAME_LOCATION_ETC = format_water_class(
            USE_CLASS_NAME_LOCATION_ETC,
            if ("sub_class" %in% names(.)) sub_class else NA_character_
          )
        ) %>%
        dplyr::select(CRITERION_ID, USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE) %>%
        dplyr::distinct() %>%
        dplyr::filter(is.finite(CRITERION_VALUE))

      # Hardness annotation (metals / range mode)
      hardness_note <- NULL
      if (!is.null(params$hardness_mode) &&
          params$hardness_mode == "range" && is.finite(rwc_val)) {
        sub_sd_metal <- sub_rp %>%
          dplyr::filter(CRITERION_ID %in% metals_ids,
                        USE_CLASS_NAME_LOCATION_ETC == "class SD waters")
        if (nrow(sub_sd_metal) > 0) {
          crit  <- sub_sd_metal$CRITERION_ID[1]
          curve <- metal_limit_curve(crit)
          if (!is.null(curve) && nrow(curve) > 0) {
            h_row <- curve %>%
              dplyr::mutate(diff = abs(limit - rwc_val)) %>%
              dplyr::slice_min(diff, n = 1, with_ties = FALSE)
            if (nrow(h_row) == 1 && is.finite(h_row$hardness)) {
              hardness_note <- paste0("RWC > WQS at Hardness values < ",
                                      round(h_row$hardness, 0),
                                      " mg/L as CaCO\u2083")
            }
          }
        }
      }

      # RP breakdown table
      cat("**RP Breakdown by Water Class**\n\n")
      rp_breakdown <- rp_df %>%
        dplyr::filter(NPDES_Pollutant == poll, perm_feature_nmbr == of) %>%
        { if (!is.null(wqs_annotation_df))
            dplyr::left_join(., wqs_annotation_df, by = "CRITERION_ID") else . } %>%
        # Same combined-display pattern as the main Findings table and the
        # plot legend — so the per-parameter table also reads
        # "class SD waters — stream" / "— reservoir/lake" where applicable.
        dplyr::mutate(
          USE_CLASS_NAME_LOCATION_ETC = format_water_class(
            USE_CLASS_NAME_LOCATION_ETC,
            if ("sub_class" %in% names(.)) sub_class else NA_character_
          )
        ) %>%
        dplyr::select(
          CRITERION_ID,
          USE_CLASS_NAME_LOCATION_ETC,
          dplyr::any_of(c("CRITERIATYPEAQUAHUMHLTH", "CRITERIATYPEFRESHSALTWATER")),
          CRITERION_VALUE,
          RP
        ) %>%
        dplyr::distinct()

      if (nrow(rp_breakdown) > 0) {
        has_type_cols <- all(c("CRITERIATYPEAQUAHUMHLTH",
                               "CRITERIATYPEFRESHSALTWATER") %in% names(rp_breakdown))
        col_names <- if (has_type_cols) {
          c("WQS ID", "Water Class", "Human / Aquatic", "Fresh / Salt",
            "WQS Criterion", "RP")
        } else {
          c("WQS ID", "Water Class", "WQS Criterion", "RP")
        }
        tbl <- kable(rp_breakdown,
                     format    = "latex",
                     booktabs  = TRUE,
                     col.names = col_names) %>%
          kable_styling(latex_options = c("HOLD_position"))
        print(tbl)
      } else {
        cat("No WQS rows found for this pollutant/outfall.\n\n")
      }

      # ── RWC / observed value ──────────────────────────────────────────────
      # Direct-comparison rows have no multiplying factor and no dilution
      # ratio: RWC_rs holds an observed extreme. Printing the MF derivation
      # for them would assert a projection that was never performed.
      is_direct_rpt <- "rp_method" %in% names(sub_rp) &&
        isTRUE(sub_rp$rp_method[1] == "direct")

      if (is_direct_rpt) {
        obs_op <- if ("rp_operator" %in% names(sub_rp)) sub_rp$rp_operator[1] else ">"
        obs_lbl <- if (!is.na(obs_op) && obs_op == "<") "minimum" else "maximum"
        obs_val <- sub_rp %>%
          dplyr::distinct(RWC_rs) %>%
          dplyr::pull() %>%
          .[is.finite(.)]
        cat("**Assessed Value**\n\n")
        cat(sprintf(
          paste0(
            "This parameter is assessed by direct comparison. The observed ",
            "%s of %s is compared to the criterion as reported: no ",
            "multiplying factor is applied and the dilution ratio is not ",
            "used.\n\n"
          ),
          obs_lbl,
          if (length(obs_val) > 0) {
            format(signif(obs_val[1], 3), scientific = FALSE)
          } else {
            "--"
          }
        ))
      } else {

      # MF step-by-step equations
      cat("**RWC Calculation**\n\n")
      if (all(c("n_used","cv","sigma_g","p_percentile","z_p","z_target","MF") %in%
              names(sub_rp))) {
        n_val   <- sub_rp %>% dplyr::distinct(n_used)       %>% dplyr::pull() %>% .[is.finite(.)]
        cv_val  <- sub_rp %>% dplyr::distinct(cv)           %>% dplyr::pull() %>% .[is.finite(.)]
        sg_val  <- sub_rp %>% dplyr::distinct(sigma_g)      %>% dplyr::pull() %>% .[is.finite(.)]
        p_val   <- sub_rp %>% dplyr::distinct(p_percentile) %>% dplyr::pull() %>% .[is.finite(.)]
        zp_val  <- sub_rp %>% dplyr::distinct(z_p)          %>% dplyr::pull() %>% .[is.finite(.)]
        zt_val  <- sub_rp %>% dplyr::distinct(z_target)     %>% dplyr::pull() %>% .[is.finite(.)]
        mf_val  <- sub_rp %>% dplyr::distinct(MF)           %>% dplyr::pull() %>% .[is.finite(.)]
        rwc_eq  <- sub_rp %>% dplyr::distinct(RWC_rs)       %>% dplyr::pull() %>% .[is.finite(.)]
        max_val <- sub_rp %>% dplyr::distinct(max_value)    %>% dplyr::pull() %>% .[is.finite(.)]
        cl_val  <- params$confidence_level
        tgt_val <- params$target_percentile

        if (length(n_val) > 0 && length(cv_val) > 0 && length(sg_val) > 0 &&
            length(p_val) > 0 && length(zp_val) > 0 && length(zt_val) > 0 &&
            length(mf_val) > 0 && is.finite(cl_val) && is.finite(tgt_val)) {

          cat(sprintf("$$\\sigma_g = \\sqrt{\\ln(%s^2+1)} = %s$$\n",
                      format(signif(cv_val[1], 2), scientific = FALSE),
                      format(signif(sg_val[1], 2), scientific = FALSE)))
          cat(sprintf("$$p = (1 - %s)^{1/%s} = %s$$\n",
                      format(signif(cl_val, 2), scientific = FALSE),
                      format(n_val[1]),
                      format(signif(p_val[1], 2), scientific = FALSE)))
          cat(sprintf("$$z_p = %s,\\; z_{\\text{target}} = %s$$\n",
                      format(signif(zp_val[1], 2), scientific = FALSE),
                      format(signif(zt_val[1], 2), scientific = FALSE)))
          cat(sprintf(
            "$$\\mathrm{MF} = \\exp\\big((%s - %s)\\times %s\\big) = %s$$\n",
            format(signif(zt_val[1], 2), scientific = FALSE),
            format(signif(zp_val[1], 2), scientific = FALSE),
            format(signif(sg_val[1], 2), scientific = FALSE),
            format(signif(mf_val[1], 2), scientific = FALSE)))
          cat(sprintf(
            "$$\\mathrm{RWC} = \\frac{\\max(x)\\times MF}{DR} = \\frac{%s\\times%s}{%s} = %s$$\n\n",
            format(signif(max_val, 2), scientific = FALSE),
            format(signif(mf_val,  2), scientific = FALSE),
            format(signif(params$dilution_ratio, 2), scientific = FALSE),
            format(signif(rwc_eq,  2), scientific = FALSE)))
        }
      }
      } # end projected/direct branch

      # ── PLOT A: Hardness vs Limit (metals / range mode only) ────────────────
      show_range_plot <-
        (!is.null(params$hardness_mode) && params$hardness_mode == "range") &&
        any(sub_rp$CRITERION_ID %in% metals_ids &
              sub_rp$USE_CLASS_NAME_LOCATION_ETC == "class SD waters")

      if (show_range_plot) {
        crit  <- sub_rp %>%
          dplyr::filter(CRITERION_ID %in% metals_ids,
                        USE_CLASS_NAME_LOCATION_ETC == "class SD waters") %>%
          dplyr::slice(1) %>% dplyr::pull(CRITERION_ID)
        curve <- metal_limit_curve(crit)
        if (!is.null(curve) && nrow(curve) > 0) {
          ylab_txt <- sub_rp$UNIT_NAME[1]
          p_h <- ggplot2::ggplot(curve, ggplot2::aes(x = hardness, y = limit)) +
            ggplot2::geom_line(color = limit_color, linewidth = 1) +
            ggplot2::scale_y_continuous(limits = c(0, NA)) +
            ggplot2::labs(
              title = paste("Hardness vs WQS Criterion \u2014", poll, "(Class SD)"),
              x     = "Hardness (mg/L as CaCO\u2083)",
              y     = ylab_txt
            ) +
            ggplot2::theme_minimal()
          knitr::opts_current$set(fig.height = 3.2)
          print(p_h)
          cat("\n\n")
        }
      }

      # ── PLOT B: Observed vs WQS / RWC ───────────────────────────────────────
      # Drawing order: WQS lines (thicker, drawn first) → RWC line (on top).
      # WQS lines are drawn individually via a loop so each gets the correct
      # class colour and linetype from the shared palette.
      if (n_samples > 0 &&
          all(c("monitoring_period_end_date", "dmr_value_nmbr") %in% names(obs))) {

        ylab_txt         <- sub_rp$UNIT_NAME[1]
        present_classes  <- unique(wqs_lines$USE_CLASS_NAME_LOCATION_ETC)

        # y-axis bounds: include obs values, WQS lines, and RWC
        all_y_vals <- c(obs$dmr_value_nmbr, wqs_lines$CRITERION_VALUE,
                        if (is.finite(rwc_val)) rwc_val else numeric(0))
        y_min <- max(0, min(all_y_vals, na.rm = TRUE) * 0.95)

        # Mirror the app: many CWA criteria sit far below any achievable
        # quantitation limit, and bacteria counts span five or six decades by
        # nature, so a linear axis collapses the criterion line onto zero.
        # Switch to log10 when that happens. This ratio is NOT a units check --
        # reported units are reconciled upstream by get_unit_conversion(), so a
        # large ratio here is normally a real exceedance or a naturally
        # wide-range parameter, not a mg/L <-> ug/L reporting error.
        pos_obs_r  <- obs$dmr_value_nmbr[is.finite(obs$dmr_value_nmbr) &
          obs$dmr_value_nmbr > 0]
        crit_min_r <- suppressWarnings(min(
          wqs_lines$CRITERION_VALUE[is.finite(wqs_lines$CRITERION_VALUE) &
            wqs_lines$CRITERION_VALUE > 0]
        ))
        oom_ratio_r <- if (length(pos_obs_r) > 0 && is.finite(crit_min_r) &&
          crit_min_r > 0) {
          max(pos_obs_r) / crit_min_r
        } else {
          NA_real_
        }
        use_log_r <- is.finite(oom_ratio_r) && oom_ratio_r >= 100
        if (use_log_r) ylab_txt <- paste0(ylab_txt, " (log scale)")

        # Non-detects (NODI B/Q, stored as 0) and literal reported zeros are
        # dropped silently by log10. Plot them at a floor equal to the smallest
        # positive value in the series, marked with open down-triangles, and
        # disclose the substitution in the caption.
        obs$is_nd_r <- is.finite(obs$dmr_value_nmbr) & obs$dmr_value_nmbr <= 0
        nd_floor_r <- if (length(pos_obs_r) > 0) min(pos_obs_r) else NA_real_
        nd_at_floor_r <- use_log_r && any(obs$is_nd_r) && is.finite(nd_floor_r)
        obs$y_plot_r <- ifelse(
          obs$is_nd_r & nd_at_floor_r,
          nd_floor_r,
          obs$dmr_value_nmbr
        )
        nd_cap_r <- if (nd_at_floor_r) {
          paste0("Open triangles are non-detects, plotted at ",
                 signif(nd_floor_r, 3), ".")
        } else if (any(obs$is_nd_r)) {
          "Open triangles are non-detects."
        } else {
          NA_character_
        }

        p <- ggplot2::ggplot(
          obs,
          ggplot2::aes(x = monitoring_period_end_date, y = y_plot_r)
        ) +
          ggplot2::geom_line(color = "lightgrey", linetype = "dotted",
                             linewidth = 0.5)

        # Draw WQS lines first (thicker), one at a time for full palette control
        if (nrow(wqs_lines) > 0) {
          for (i in seq_len(nrow(wqs_lines))) {
            row_i <- wqs_lines[i, ]
            p <- p + ggplot2::geom_hline(
              yintercept  = row_i$CRITERION_VALUE,
              color       = wqs_colors[row_i$USE_CLASS_NAME_LOCATION_ETC],
              linetype    = wqs_linetypes[row_i$USE_CLASS_NAME_LOCATION_ETC],
              linewidth   = wqs_lwd_max,
              inherit.aes = FALSE
            )
          }
        }

        # RWC line drawn last — on top of all WQS lines
        if (is.finite(rwc_val)) {
          p <- p + ggplot2::geom_hline(
            yintercept  = rwc_val,
            color       = rwc_color,
            linetype    = "solid",
            linewidth   = rwc_lwd,
            inherit.aes = FALSE
          )
        }

        # Observation points (colour by reported unit — second colour scale)
        p <- p +
          ggplot2::geom_point(
            data = obs[!obs$is_nd_r, , drop = FALSE],
            ggplot2::aes(color = dmr_unit_desc),
            alpha = 0.85
          ) +
          ggplot2::scale_color_manual(
            name     = "Unit",
            values   = unit_colors,
            na.value = "#999999",
            breaks   = names(unit_colors),
            drop     = FALSE
          ) +
          {
            if (use_log_r) {
              ggplot2::scale_y_log10()
            } else {
              ggplot2::scale_y_continuous(limits = c(y_min, NA))
            }
          } +
          ggplot2::labs(
            x = "Monitoring date",
            y = ylab_txt,
            caption = if (!is.na(nd_cap_r)) nd_cap_r else NULL
          ) +
          ggplot2::theme_minimal() +
          ggplot2::theme(
            legend.position = "bottom",
            legend.box      = "vertical",
            legend.text     = ggplot2::element_text(size = 7),
            legend.title    = ggplot2::element_text(size = 8),
            plot.caption    = ggplot2::element_text(hjust = 0, color = rwc_color, size = 8)
          )

        # Non-detect markers drawn after the unit scale so they take no colour
        # from it. inherit.aes keeps x and y_plot_r from the base mapping.
        if (any(obs$is_nd_r)) {
          p <- p + ggplot2::geom_point(
            data  = obs[obs$is_nd_r, , drop = FALSE],
            shape = 6, color = "grey30", size = 2.2, stroke = 0.9
          )
        }

        # Water class legend via zero-alpha ghost layer (separate from unit scale)
        if (length(present_classes) > 0) {
          p <- p + ggnewscale::new_scale_color() +
            ggplot2::geom_hline(
              data = wqs_lines,
              ggplot2::aes(yintercept  = CRITERION_VALUE,
                           color       = USE_CLASS_NAME_LOCATION_ETC,
                           linetype    = USE_CLASS_NAME_LOCATION_ETC),
              linewidth   = wqs_lwd_max,
              inherit.aes = FALSE,
              show.legend = TRUE
            ) +
            ggplot2::scale_color_manual(
              name   = "Water Class",
              values = active_wqs_colors(present_classes)
            ) +
            ggplot2::scale_linetype_manual(
              name   = "Water Class",
              values = active_wqs_linetypes(present_classes)
            )
        }

        # Hardness note goes in the caption (below the plot) rather than as an
        # in-panel annotation, which previously ran off the right page margin.
        if (!is.null(hardness_note) && is.finite(rwc_val)) {
          p <- p + ggplot2::labs(caption = hardness_note)
        }

        knitr::opts_current$set(fig.height = 3.4)
        print(p)
        cat("\n\n")

      } else {
        cat("No dated observations to plot.\n\n")
      }

      # Detail table (WQS info per criterion). Mirrors rp_breakdown above
      # for the post-plot table that also shows the calculated RWC.
      detail <- sub_rp %>%
        { if (!is.null(wqs_annotation_df))
            dplyr::left_join(., wqs_annotation_df, by = "CRITERION_ID") else . } %>%
        dplyr::mutate(
          USE_CLASS_NAME_LOCATION_ETC = format_water_class(
            USE_CLASS_NAME_LOCATION_ETC,
            if ("sub_class" %in% names(.)) sub_class else NA_character_
          )
        ) %>%
        dplyr::select(
          CRITERION_ID,
          USE_CLASS_NAME_LOCATION_ETC,
          dplyr::any_of(c("CRITERIATYPEAQUAHUMHLTH", "CRITERIATYPEFRESHSALTWATER")),
          CRITERION_VALUE,
          RWC_rs,
          RP
        ) %>%
        dplyr::distinct()

      if (nrow(detail) > 0) {
        has_type_cols <- all(c("CRITERIATYPEAQUAHUMHLTH",
                               "CRITERIATYPEFRESHSALTWATER") %in% names(detail))
        col_names <- if (has_type_cols) {
          c("WQS ID", "Water Class", "Human Health Type", "Fresh/Salt Type",
            "WQS Criterion", "RWC", "RP")
        } else {
          c("WQS ID", "Water Class", "WQS Criterion", "RWC", "RP")
        }
        kable(detail,
              format    = "latex",
              booktabs  = TRUE,
              col.names = col_names) %>%
          kable_styling(latex_options = c("HOLD_position")) %>%
          print()
      }

    } # end outfall loop
  } # end pollutant loop
} # end dmr_df check
```


# Unavailable Pollutants {#sec-unavailable}

```{r}
#| label: unavailable
#| echo: false

# Rebuild bottom here so this chunk is self-contained and does not depend on
# the coverage chunk having run successfully (e.g. in standalone mode where
# coverage_df may be NULL).
bottom_unavail <- if (!is.null(coverage_df) && nrow(coverage_df) > 0) {
  coverage_df %>%
    dplyr::filter(`# Samples` == 0) %>%
    dplyr::arrange(Pollutant) %>%
    dplyr::select(!dplyr::any_of(c("Form", "User Data"))) %>%
    dplyr::distinct()
} else {
  NULL
}

if (!is.null(bottom_unavail) && nrow(bottom_unavail) > 0) {
  kable(bottom_unavail,
        format    = "latex",
        booktabs  = TRUE,
        longtable = TRUE,
        col.names = c("Pollutant", "Ref Param", "# Samples")) %>%
    kable_styling(latex_options = c("striped", "repeat_header"))
} else {
  cat("All expected pollutants had data available.")
}
```




# Excluded Parameters {#sec-excluded}

```{r}
#| label: excluded-params
#| echo: false
#| results: asis

excl_df <- if (!is.null(overrides_df) && nrow(overrides_df) > 0) {
  overrides_df %>%
    dplyr::filter(decision == "exclude") %>%
    dplyr::select(
      `Parameter`      = parameter_desc,
      `Code`           = parameter_code,
      `Flag Reason`    = case_reason,
      `Samples in DMR` = dplyr::any_of("n_samples"),
      `Reason for Exclusion` = notes
    )
} else NULL

if (!is.null(excl_df) && nrow(excl_df) > 0) {
  cat("The following parameters had DMR data available but were excluded from the RP ",
      "analysis at the direction of the user.\n\n")
  kable(excl_df,
        format    = "latex",
        booktabs  = TRUE,
        longtable = TRUE) %>%
    kable_styling(latex_options = c("striped", "repeat_header")) %>%
    column_spec(1, width = "3.5cm") %>%   # Parameter name
    column_spec(5, width = "5cm")          # Reason for Exclusion — wraps long auto-exclude text
} else {
  cat("No parameters were explicitly excluded from this analysis.")
}
```


# Manual WQS Overrides {#sec-manual-wqs}

```{r}
#| label: manual-wqs
#| echo: false
#| results: asis

manual_df <- if (!is.null(overrides_df) && nrow(overrides_df) > 0) {
  overrides_df %>%
    dplyr::filter(decision == "include") %>%
    dplyr::select(
      `Parameter`        = parameter_desc,
      `Code`             = parameter_code,
      `Flag Reason`      = case_reason,
      `Water Class(es)`  = water_classes,
      `Criterion Value`  = criterion_value,
      `Unit`             = unit,
      `Criterion Type`   = criterion_type,
      `Basis / Notes`    = notes
    )
} else NULL

if (!is.null(manual_df) && nrow(manual_df) > 0) {
  cat(paste0(
    "The following parameters had no matching WQS criterion for the selected water ",
    "class and were evaluated using manually entered criteria. These criteria were ",
    "supplied by the permit writer and are documented here as part of the permanent ",
    "record. All manually entered criteria are flagged with a dagger symbol in the Coverage ",
    "Summary table.\n\n"
  ))
  # print() is explicit: under results:asis an auto-printed kable can be
  # skipped, so force the LaTeX out.
  print(
  kable(manual_df,
        format    = "latex",
        booktabs  = TRUE,
        longtable = TRUE,
        linesep   = "",
        # Center the three short columns: Water Class(es), Criterion Value,
        # Criterion Type. align applies to both header and body cells.
        align     = c("l", "l", "l", "c", "c", "l", "c", "l")) %>%
    kable_styling(latex_options = c("striped", "repeat_header")) %>%
    # Fixed widths force wrapping at full font size. LaTeX adds ~0.42cm of
    # padding (2\\tabcolsep) PER column on top of these p{} widths, so eight
    # columns carry ~3.4cm of padding; the content widths below sum to 11.9cm,
    # for ~15.3cm total inside the 16.5cm text block. Parameter and
    # Basis/Notes are the narrow-and-wrapping columns per request.
    column_spec(1, width = "2.4cm") %>%   # Parameter (wraps)
    column_spec(2, width = "1.0cm") %>%   # Code
    column_spec(3, width = "1.6cm") %>%   # Flag Reason (wraps)
    column_spec(4, width = "1.7cm") %>%   # Water Class(es) (wraps, centered)
    column_spec(5, width = "1.5cm") %>%   # Criterion Value (centered)
    column_spec(6, width = "0.8cm") %>%   # Unit
    column_spec(7, width = "1.3cm") %>%   # Criterion Type (centered)
    column_spec(8, width = "1.9cm")       # Basis / Notes (wraps)
  )
} else {
  cat("No manual WQS criteria were entered for this analysis.")
}
```
