---
title: "NPDES Reasonable Potential (RP) Report"
subtitle: "`r paste('Report Generated:', format(Sys.Date(), '%B %d, %Y'))`"
format:
  pdf:
    toc: true
    toc-depth: 1
    number-sections: true
    geometry: margin=1in
execute:
  echo: false
params:
  permit_id: ""
  facility_name: ""
  date_start: ""
  date_end: ""
  forms: ""
  dilution_ratio: 1
  confidence_level: 0.95
  target_percentile: 0.95
  hardness_mode: "range"
  hardness_value: NA
  rp_path: ""
  coverage_path: ""
  dmr_path: ""
  ph_dmr_path: ""
  temp_dmr_path: ""
  wqs_info_path: ""   # <-- add
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, out.width = "\\linewidth", fig.align = "center", fig.show = "asis")

library(readr)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(lubridate)
library(ggforce)
library(ggnewscale)
library(tidyr)

rp_df       <- readr::read_csv(params$rp_path, show_col_types = FALSE)
coverage_df <- if (nzchar(params$coverage_path)) readr::read_csv(params$coverage_path, show_col_types = FALSE) else NULL
dmr_df      <- if (nzchar(params$dmr_path))      readr::read_csv(params$dmr_path,      show_col_types = FALSE) else NULL
wqs_info_df <- if (nzchar(params$wqs_info_path)) {
  readr::read_csv(params$wqs_info_path, show_col_types = FALSE) %>%
    dplyr::select(-USE_CLASS_NAME_LOCATION_ETC)
} else NULL

# Count outfalls (unique perm_feature_nmbr)
outfall_count <- if (!is.null(dmr_df) && "perm_feature_nmbr" %in% names(dmr_df)) {
  dplyr::n_distinct(dmr_df$perm_feature_nmbr)
} else {
  0L
}

# Permit limit history — guarded for standalone runs where ECHO limit columns may be absent
limits_required_cols <- c("parameter_code", "limit_value_nmbr", "limit_unit_desc",
                           "limit_begin_date", "limit_end_date")
limits <- if (!is.null(dmr_df) && all(limits_required_cols %in% names(dmr_df))) {
  dmr_df %>%
    dplyr::select(dplyr::all_of(limits_required_cols)) %>%
    tidyr::drop_na() %>%
    dplyr::mutate(
      limit_begin_date = lubridate::mdy(limit_begin_date),
      limit_end_date   = lubridate::mdy(limit_end_date)
    ) %>%
    dplyr::distinct()
} else {
  NULL
}

# Metal criterion IDs and constants (for on-the-fly hardness curves in the report)
metals_ids <- c("79228","79238","79236","79253","79248","79251","79264")

metal_constants <- data.frame(
  CRITERION_ID = metals_ids,
  LC = c(0.7977, 0.8190, 0.8545, 1.273, 0.8460, 1.72, 0.8473),
  RC = c(3.909, 0.6848, 1.702, 4.705, 0.0584, 6.59, 0.884)
)

# Build curve of limits for a given CRITERION_ID across hardness 25..400 (step = 1)
metal_limit_curve <- function(crit_id, h_min = 25, h_max = 400) {
  row <- metal_constants[metal_constants$CRITERION_ID == crit_id, ]
  if (nrow(row) == 0) return(NULL)
  hardness <- seq(h_min, h_max, 1)
  limit <- exp(row$LC * log(hardness) - row$RC)
  dplyr::tibble(CRITERION_ID = crit_id, hardness = hardness, limit = limit)
}

# Safe readers for optional inputs
safe_read_csv <- function(path) {
  if (is.character(path) && nzchar(path) && file.exists(path)) {
    readr::read_csv(path, show_col_types = FALSE)
  } else {
    NULL
  }
}

ph_dmr_df   <- safe_read_csv(params$ph_dmr_path)
temp_dmr_df <- safe_read_csv(params$temp_dmr_path)

# Normalize dates if present
if (is.data.frame(ph_dmr_df) && "monitoring_period_end_date" %in% names(ph_dmr_df)) {
  if (!inherits(ph_dmr_df$monitoring_period_end_date, "Date")) {
    ph_dmr_df$monitoring_period_end_date <- suppressWarnings(lubridate::mdy(ph_dmr_df$monitoring_period_end_date))
  }
}
if (is.data.frame(temp_dmr_df) && "monitoring_period_end_date" %in% names(temp_dmr_df)) {
  if (!inherits(temp_dmr_df$monitoring_period_end_date, "Date")) {
    temp_dmr_df$monitoring_period_end_date <- suppressWarnings(lubridate::mdy(temp_dmr_df$monitoring_period_end_date))
  }
}

# ── Shared visual palette ────────────────────────────────────────────────────
# Single source of truth for water class colors and linetypes.
# Used identically in pH, temperature, and concentration-based plots so the
# legend is immediately recognisable across all sections of the report.
#
# Drawing order rule: WQS lines are drawn thickest → thinnest so no line is
# buried under a wider one.  Within WQS lines, Min bounds (lwr_lwd) are drawn
# before Max bounds (upr_lwd); the RWC line is drawn last (on top).
#
wqs_colors <- c(
  "class SB waters"                 = "#1f78b4",  # blue
  "class SD waters"                 = "#ff7f00",  # orange
  "class SD waters- drinking water" = "#33a02c",  # green
  "class SG waters"                 = "#6a3d9a",  # purple
  "class SG waters- drinking water" = "#e31a1c",  # red
  "surface waters"                  = "#b15928"   # brown
)

wqs_linetypes <- c(
  "class SB waters"                 = "dashed",
  "class SD waters"                 = "dashed",
  "class SD waters- drinking water" = "dotdash",
  "class SG waters"                 = "dashed",
  "class SG waters- drinking water" = "dotdash",
  "surface waters"                  = "longdash"
)

# Line widths: Max-bound lines are thinner and drawn on top; Min-bound lines
# are thicker and drawn first so Max is always visible above them.
wqs_lwd_max <- 0.7   # upper limit lines (Max pH, or WQS criterion for conc.)
wqs_lwd_min <- 1.3   # lower limit lines (Min pH)
rwc_color   <- "#2c7fb8"
rwc_lwd     <- 1.0   # RWC line — drawn last, distinct colour keeps it visible
limit_color <- "#d61309"  # hardness curve line colour

# Helper: given a vector of class names present in a plot, return only the
# palette entries that are needed (avoids orphan legend keys).
active_wqs_colors    <- function(classes) wqs_colors[intersect(names(wqs_colors), classes)]
active_wqs_linetypes <- function(classes) wqs_linetypes[intersect(names(wqs_linetypes), classes)]

# Preset point colors by reported unit (dmr_unit_desc) — unchanged from before
unit_levels <- c(
  "deg C","mg/L","SU","mL/L","ug/L","MGD","NTU","col unit (pc)",
  "MF/L","kg/d","#/100mL","%","d","gal/d","MPN/100mL","m3/d"
)
unit_colors <- setNames(
  c(
    "#1b9e77", "#d95f02", "#7570b3", "#e7298a",
    "#66a61e", "#e6ab02", "#a6761d", "#1f78b4",
    "#b2df8a", "#fb9a99", "#e31a1c", "#ff7f00",
    "#6a3d9a", "#b15928", "#33a02c", "#a6cee3"
  ),
  unit_levels
)
```

# Summary

Permit: `r params$permit_id`  
Facility: `r params$facility_name`  
Date range: `r params$date_start` to `r params$date_end`  
Selected forms: `r params$forms`  
Dilution ratio: `r params$dilution_ratio`  
Confidence level: `r params$confidence_level`  
Target percentile: `r params$target_percentile`  
Hardness setting: `r if (params$hardness_mode == "set") paste0("Set (", params$hardness_value, " mg/L as CaCO3)") else "Range"`  

## Included Pollutants

The following table shows the available data for each pollutant listed in the selected NPDES application forms:(`r params$forms`). For a complete list of pollutants that had no associated data, refer to  @sec-unavailable

```{r}
#| label: coverage
#| echo: false

if (!is.null(coverage_df) && nrow(coverage_df) > 0) {
  
  # Reorder so included are at the top
  top <- coverage_df%>%
    filter(`# Samples` > 0)%>%
    arrange(Pollutant)%>%
    select(!Form)%>%
    distinct()
  bottom <- coverage_df%>%
    filter(`# Samples` == 0)%>%
    arrange(Pollutant)%>%
    select(!Form)%>%
    distinct()
  
  #coverage_reorder <- rbind(top,bottom)
  
  
  kable(top, 
      format = "latex", 
      booktabs = TRUE, 
      longtable = TRUE, # Allows the table to span multiple pages
      col.names = c("Pollutant", "Ref Param", "# Samples")) %>%
  kable_styling(latex_options = c("striped", "repeat_header")) # Repeat header on new pages
  # row_spec(nrow(top), 
  #        extra_latex_after = "\\arrayrulecolor{red}\\midrule\\arrayrulecolor{black}")
 
    
} else {
  cat("No coverage summary available.")
}
```

# RP Findings

RP column: YES indicates RWC exceeds the criterion (or pH/temperature logic).  
RWC values rounded to 2 decimals in the app table.

```{r}
#| label: rp-table
#| echo: false

rp_df %>%
  mutate(RWC_rs = round(RWC_rs, 2),
         CRITERION_VALUE = round(CRITERION_VALUE,2)) %>%
  drop_na(RWC_rs)%>%
  select(CRITERION_ID, NPDES_Pollutant,perm_feature_nmbr, USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE, RWC_rs, RP) %>%
  arrange(desc(RP), NPDES_Pollutant) %>%
  distinct()%>%
  kable(col.names = c("WQS ID","Pollutant","Outfall","Water Class","Limit","RWC","RP"))
```

# Non-Concentration Based Measures

```{r}
#| label: receiving-water
#| echo: false
#| results: asis

# Helper for existence checks
has_rows <- function(df) is.data.frame(df) && nrow(df) > 0

if (is.null(rp_df)) {
  cat("No RP results are available; skipping receiving water sections.\n\n")
} else {
  # Build outfall list from whatever is available
  outfalls <- unique(c(
    if (has_rows(ph_dmr_df))   ph_dmr_df$perm_feature_nmbr else character(0),
    if (has_rows(temp_dmr_df)) temp_dmr_df$perm_feature_nmbr else character(0),
    if (has_rows(rp_df))       rp_df$perm_feature_nmbr else character(0)
  ))
  outfalls <- sort(unique(outfalls[!is.na(outfalls)]))

  if (length(outfalls) == 0 || (!has_rows(ph_dmr_df) && !has_rows(temp_dmr_df))) {
    cat("No pH or temperature data to report.\n\n")
  } else {

    for (of in outfalls) {

      # ── pH ──────────────────────────────────────────────────────────────────
      if (has_rows(ph_dmr_df)) {
        ph_obs <- ph_dmr_df %>%
          dplyr::filter(perm_feature_nmbr == of) %>%
          dplyr::arrange(monitoring_period_end_date)

        if (nrow(ph_obs) > 0) {
          cat(sprintf("## Outfall %s — pH\n\n", of))

          # Summary table
          ph_summ <- ph_obs %>%
            dplyr::summarise(
              max_pH = max(dmr_value_nmbr, na.rm = TRUE),
              min_pH = min(dmr_value_nmbr, na.rm = TRUE)
            )
          print(knitr::kable(ph_summ, format = "latex", booktabs = TRUE,
                             col.names = c("Max pH", "Min pH")))
          cat("\n\n")

          # WQS lines from rp_df for pH at this outfall.
          # Each row has USE_CLASS_NAME_LOCATION_ETC + Method (Max / Min).
          # We build a data frame with one row per line, tagged so we can
          # control drawing order: Min lines first (thick), Max lines on top (thin).
          wqs_ph_raw <- rp_df %>%
            dplyr::filter(parameter_code == "00400", perm_feature_nmbr == of) %>%
            dplyr::select(USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE, Method) %>%
            dplyr::distinct() %>%
            dplyr::filter(is.finite(CRITERION_VALUE))

          # Build combined line-key used for colour + linetype scale lookup
          wqs_ph <- wqs_ph_raw %>%
            dplyr::mutate(
              line_key = USE_CLASS_NAME_LOCATION_ETC,
              # linewidth: Min bound drawn thicker first, Max drawn thinner on top
              lwd = dplyr::if_else(Method == "Min", wqs_lwd_min, wqs_lwd_max),
              # draw order: Min rows first so Max is rendered on top
              draw_order = dplyr::if_else(Method == "Min", 1L, 2L)
            ) %>%
            dplyr::arrange(draw_order)

          present_classes <- unique(wqs_ph$line_key)

          # y-axis lower bound: a unit below the smallest value among obs + WQS
          min_value <- min(c(ph_obs$dmr_value_nmbr, wqs_ph$CRITERION_VALUE),
                          na.rm = TRUE) - 1

          p <- ggplot2::ggplot(
            ph_obs,
            ggplot2::aes(x = monitoring_period_end_date, y = dmr_value_nmbr)
          ) +
            ggplot2::geom_line(color = "lightgrey", linetype = "dotted",
                               linewidth = 0.5)

          # Draw WQS lines in draw_order (Min first, Max on top).
          # Each row gets its own linewidth via a loop so we can vary lwd
          # per row — geom_hline does not support a lwd aesthetic directly.
          if (nrow(wqs_ph) > 0) {
            for (i in seq_len(nrow(wqs_ph))) {
              row_i <- wqs_ph[i, ]
              p <- p + ggplot2::geom_hline(
                yintercept = row_i$CRITERION_VALUE,
                color      = wqs_colors[row_i$line_key],
                linetype   = wqs_linetypes[row_i$line_key],
                linewidth  = row_i$lwd,
                inherit.aes = FALSE
              )
            }
          }

          p <- p +
            ggplot2::geom_point(size = 1.8, color = "#7570b3") +
            ggplot2::scale_y_continuous(limits = c(min_value, NA)) +
            ggplot2::labs(
              title    = sprintf("pH over time — Outfall %s", of),
              x        = "Date",
              y        = "pH (SU)"
            ) +
            ggplot2::theme_minimal()

          # Add a manual legend for water class lines.
          # We show Max lines in the legend (representative of each class).
          if (length(present_classes) > 0) {
            legend_df <- wqs_ph %>%
              dplyr::filter(draw_order == 2L) %>%   # Max rows only for legend
              dplyr::distinct(line_key, .keep_all = TRUE)

            p <- p +
              ggplot2::geom_hline(
                data        = legend_df,
                ggplot2::aes(yintercept = CRITERION_VALUE,
                             color      = line_key,
                             linetype   = line_key),
                linewidth   = wqs_lwd_max,
                inherit.aes = FALSE,
                show.legend = TRUE
              ) +
              ggplot2::scale_color_manual(
                name   = "Water Class",
                values = active_wqs_colors(present_classes)
              ) +
              ggplot2::scale_linetype_manual(
                name   = "Water Class",
                values = active_wqs_linetypes(present_classes)
              )
          }

          knitr::opts_current$set(out.width = "\\linewidth",
                                  fig.align = "center", fig.show = "asis")
          print(p)
          cat("\n\n")

          # RP table under pH plot
          ph_rp_tbl <- wqs_ph_raw %>%
            dplyr::left_join(
              rp_df %>%
                dplyr::filter(parameter_code == "00400", perm_feature_nmbr == of) %>%
                dplyr::select(USE_CLASS_NAME_LOCATION_ETC, CRITERION_ID, Method, RP) %>%
                dplyr::distinct(),
              by = c("USE_CLASS_NAME_LOCATION_ETC", "Method")
            ) %>%
            dplyr::mutate(
              Observed = dplyr::if_else(
                Method == "Max",
                round(max(ph_obs$dmr_value_nmbr, na.rm = TRUE), 2),
                round(min(ph_obs$dmr_value_nmbr, na.rm = TRUE), 2)
              ),
              CRITERION_VALUE = round(CRITERION_VALUE, 2)
            ) %>%
            dplyr::select(CRITERION_ID, USE_CLASS_NAME_LOCATION_ETC,
                          Method, CRITERION_VALUE, Observed, RP)

          if (nrow(ph_rp_tbl) > 0) {
            tbl <- knitr::kable(
              ph_rp_tbl,
              format    = "latex",
              booktabs  = TRUE,
              col.names = c("WQS ID", "Water Class", "Bound",
                            "WQS Limit", "Observed pH", "RP")
            ) %>% kableExtra::kable_styling(latex_options = c("hold_position"))
            print(tbl)
            cat("\n\n")
          }
        }
      } # end pH

      # ── Temperature ─────────────────────────────────────────────────────────
      if (has_rows(temp_dmr_df)) {
        t_obs <- temp_dmr_df %>%
          dplyr::filter(perm_feature_nmbr == of) %>%
          dplyr::arrange(monitoring_period_end_date)

        if (nrow(t_obs) > 0) {
          # Use a sub-heading if pH was also present for this outfall
          if (has_rows(ph_dmr_df) &&
              nrow(dplyr::filter(ph_dmr_df, perm_feature_nmbr == of)) > 0) {
            cat(sprintf("### Temperature — Outfall %s\n\n", of))
          } else {
            cat(sprintf("## Outfall %s — Temperature\n\n", of))
          }

          # Summary table
          t_summ <- t_obs %>%
            dplyr::summarise(Max_Temperature_C = max(dmr_value_nmbr, na.rm = TRUE))
          print(knitr::kable(t_summ, format = "latex", booktabs = TRUE,
                             col.names = c("Max Temperature (°C)")))
          cat("\n\n")

          # WQS lines for temperature — temperature is a single upper bound per
          # class (no Min/Max pair), so no linewidth split needed.
          # Still ordered so thicker lines are drawn first.
          wqs_t <- rp_df %>%
            dplyr::filter(parameter_code == "00010", perm_feature_nmbr == of) %>%
            dplyr::select(USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE) %>%
            dplyr::distinct() %>%
            dplyr::filter(is.finite(CRITERION_VALUE)) %>%
            # Draw thicker lines first (more classes = draw in consistent order)
            dplyr::arrange(USE_CLASS_NAME_LOCATION_ETC)

          present_t_classes <- unique(wqs_t$USE_CLASS_NAME_LOCATION_ETC)

          p2 <- ggplot2::ggplot(
            t_obs,
            ggplot2::aes(x = monitoring_period_end_date, y = dmr_value_nmbr)
          ) +
            ggplot2::geom_line(color = "lightgrey", linetype = "dotted",
                               linewidth = 0.5)

          # Draw WQS lines individually to use consistent lwd
          if (nrow(wqs_t) > 0) {
            for (i in seq_len(nrow(wqs_t))) {
              row_i <- wqs_t[i, ]
              p2 <- p2 + ggplot2::geom_hline(
                yintercept  = row_i$CRITERION_VALUE,
                color       = wqs_colors[row_i$USE_CLASS_NAME_LOCATION_ETC],
                linetype    = wqs_linetypes[row_i$USE_CLASS_NAME_LOCATION_ETC],
                linewidth   = wqs_lwd_max,
                inherit.aes = FALSE
              )
            }
          }

          p2 <- p2 +
            ggplot2::geom_point(size = 1.8, color = "#1b9e77") +
            ggplot2::scale_y_continuous(limits = c(0, NA)) +
            ggplot2::labs(
              title = sprintf("Temperature over time — Outfall %s", of),
              x     = "Date",
              y     = "Temperature (°C)"
            ) +
            ggplot2::theme_minimal()

          # Legend via zero-alpha ghost layer
          if (length(present_t_classes) > 0) {
            p2 <- p2 +
              ggplot2::geom_hline(
                data        = wqs_t,
                ggplot2::aes(yintercept  = CRITERION_VALUE,
                             color       = USE_CLASS_NAME_LOCATION_ETC,
                             linetype    = USE_CLASS_NAME_LOCATION_ETC),
                linewidth   = wqs_lwd_max,
                inherit.aes = FALSE,
                show.legend = TRUE
              ) +
              ggplot2::scale_color_manual(
                name   = "Water Class",
                values = active_wqs_colors(present_t_classes)
              ) +
              ggplot2::scale_linetype_manual(
                name   = "Water Class",
                values = active_wqs_linetypes(present_t_classes)
              )
          }

          knitr::opts_current$set(out.width = "\\linewidth",
                                  fig.align = "center", fig.show = "asis")
          print(p2)
          cat("\n\n")

          # Temperature RP table
          temp_rp_tbl <- rp_df %>%
            dplyr::filter(parameter_code == "00010", perm_feature_nmbr == of) %>%
            dplyr::select(CRITERION_ID, USE_CLASS_NAME_LOCATION_ETC,
                          CRITERION_VALUE, RP) %>%
            dplyr::distinct() %>%
            dplyr::mutate(
              Observed_Max = round(max(t_obs$dmr_value_nmbr, na.rm = TRUE), 2),
              CRITERION_VALUE = round(CRITERION_VALUE, 2)
            )

          if (nrow(temp_rp_tbl) > 0) {
            tbl <- knitr::kable(
              temp_rp_tbl,
              format    = "latex",
              booktabs  = TRUE,
              col.names = c("WQS ID", "Water Class", "Limit (°C)",
                            "RP", "Observed Max (°C)")
            ) %>% kableExtra::kable_styling(latex_options = c("hold_position"))
            print(tbl)
            cat("\n\n")
          }
        }
      } # end temperature

    } # end outfall loop
  }

} # end rp_df is not NULL
```




# Concentration Based Measures

## RP Equations

For concentration based measures, the following equations are used to estimate Receiving Water Concentration (RWC)

$$\sigma_g = \sqrt{\ln(\mathrm{CV}^2 + 1)}$$
$$p = (1 - \mathrm{CL})^{1/n}$$
$$z_p = \Phi^{-1}(p),\; z_{\text{target}} = \Phi^{-1}(\mathrm{target})$$
$$\mathrm{MF} = \exp\big((z_{\text{target}} - z_p)\,\sigma_g\big)$$

$$\text{RWC} = \frac{\max(x) \times MF}{DR}$$

```{r}
#| label: per-pollutant
#| echo: false
#| results: asis

if (is.null(dmr_df) || nrow(dmr_df) == 0) {
  cat("No DMR data were provided; per-pollutant sections are omitted.")
} else {
  pollutants <- rp_df %>%
    dplyr::filter(RWC_rs > 0) %>%
    dplyr::distinct(NPDES_Pollutant) %>%
    dplyr::pull()

  for (poll in pollutants) {
    sub_all <- rp_df %>% dplyr::filter(NPDES_Pollutant == poll)
    param_codes <- sub_all %>% dplyr::distinct(parameter_code) %>% dplyr::pull()
    param_code <- if (length(param_codes) > 0) param_codes[1] else NA_character_

    cat(sprintf("\n\n## %s\n\n", poll))

    # Permit limit history (guarded — absent in standalone runs)
    if (!is.null(limits) && is.data.frame(limits) && !is.na(param_code)) {
      lim_tbl <- limits %>% dplyr::filter(parameter_code == param_code)
      if (nrow(lim_tbl) > 0) {
        lim_tbl_disp <- lim_tbl %>%
          dplyr::transmute(
            `Limit Value` = limit_value_nmbr,
            `Unit`        = limit_unit_desc,
            `Limit Start` = format(limit_begin_date, "%m/%d/%Y"),
            `Limit End`   = format(limit_end_date,   "%m/%d/%Y")
          ) %>%
          dplyr::distinct()
        cat("**Permit Limit History**\n\n")
        kable(lim_tbl_disp,
              format    = "latex",
              booktabs  = TRUE,
              col.names = c("Limit Value", "Unit", "Limit Start", "Limit End")) %>%
          kable_styling(latex_options = c("hold_position")) %>%
          print()
      }
    }

    outfalls_for_poll <- if ("perm_feature_nmbr" %in% names(sub_all)) {
      unique(sub_all$perm_feature_nmbr)
    } else if (!is.null(dmr_df) && "perm_feature_nmbr" %in% names(dmr_df)) {
      unique(dplyr::filter(dmr_df, parameter_code == param_code)$perm_feature_nmbr)
    } else NA_character_
    outfalls_for_poll <- outfalls_for_poll[!is.na(outfalls_for_poll)]
    if (length(outfalls_for_poll) == 0) outfalls_for_poll <- NA_character_

    show_outfall_subheads <- is.numeric(outfall_count) && outfall_count > 1

    for (of in outfalls_for_poll) {
      if (show_outfall_subheads && !is.na(of)) cat(sprintf("### Outfall %s\n\n", of))

      sub_rp <- if (!is.na(of)) sub_all %>% dplyr::filter(perm_feature_nmbr == of) else sub_all

      obs <- dmr_df %>%
        dplyr::filter(parameter_code == param_code) %>%
        { if (!is.na(of) && "perm_feature_nmbr" %in% names(.))
            dplyr::filter(., perm_feature_nmbr == of) else . } %>%
        dplyr::mutate(
          monitoring_period_end_date = lubridate::mdy(monitoring_period_end_date)
        ) %>%
        dplyr::arrange(monitoring_period_end_date)

      n_samples <- nrow(obs)
      unit_conv_flag <- if ("Conv_Flag" %in% names(obs))
        any(obs$Conv_Flag == "PASS", na.rm = TRUE) else NA
      cat(sprintf("- Samples: %d\n\n", n_samples))
      if (!is.na(unit_conv_flag))
        cat(sprintf("- Units converted: %s\n\n",
                    if (unit_conv_flag) "Yes" else "No"))

      # RWC and WQS lines for this outfall
      rwc_val <- sub_rp %>%
        dplyr::distinct(RWC_rs) %>% dplyr::pull(RWC_rs) %>% .[is.finite(.)]
      rwc_val <- if (length(rwc_val) > 0) rwc_val[1] else NA_real_

      wqs_lines <- sub_rp %>%
        dplyr::select(CRITERION_ID, USE_CLASS_NAME_LOCATION_ETC, CRITERION_VALUE) %>%
        dplyr::distinct() %>%
        dplyr::filter(is.finite(CRITERION_VALUE))

      # Hardness annotation (metals / range mode)
      hardness_note <- NULL
      if (!is.null(params$hardness_mode) &&
          params$hardness_mode == "range" && is.finite(rwc_val)) {
        sub_sd_metal <- sub_rp %>%
          dplyr::filter(CRITERION_ID %in% metals_ids,
                        USE_CLASS_NAME_LOCATION_ETC == "class SD waters")
        if (nrow(sub_sd_metal) > 0) {
          crit  <- sub_sd_metal$CRITERION_ID[1]
          curve <- metal_limit_curve(crit)
          if (!is.null(curve) && nrow(curve) > 0) {
            h_row <- curve %>%
              dplyr::mutate(diff = abs(limit - rwc_val)) %>%
              dplyr::slice_min(diff, n = 1, with_ties = FALSE)
            if (nrow(h_row) == 1 && is.finite(h_row$hardness)) {
              hardness_note <- paste0("RWC > WQS at Hardness values < ",
                                      round(h_row$hardness, 0),
                                      " mg/L as CaCO\u2083")
            }
          }
        }
      }

      # RP breakdown table
      cat("**RP Breakdown by Water Class**\n\n")
      rp_breakdown <- rp_df %>%
        dplyr::filter(NPDES_Pollutant == poll, perm_feature_nmbr == of) %>%
        { if (!is.null(wqs_info_df))
            dplyr::left_join(., wqs_info_df, by = "CRITERION_ID") else . } %>%
        dplyr::select(
          CRITERION_ID,
          USE_CLASS_NAME_LOCATION_ETC,
          dplyr::any_of(c("CRITERIATYPEAQUAHUMHLTH", "CRITERIATYPEFRESHSALTWATER")),
          CRITERION_VALUE,
          RP
        ) %>%
        dplyr::distinct()

      if (nrow(rp_breakdown) > 0) {
        has_type_cols <- all(c("CRITERIATYPEAQUAHUMHLTH",
                               "CRITERIATYPEFRESHSALTWATER") %in% names(rp_breakdown))
        col_names <- if (has_type_cols) {
          c("WQS ID", "Water Class", "Human / Aquatic", "Fresh / Salt",
            "WQS Limit", "RP")
        } else {
          c("WQS ID", "Water Class", "WQS Limit", "RP")
        }
        tbl <- kable(rp_breakdown,
                     format    = "latex",
                     booktabs  = TRUE,
                     col.names = col_names) %>%
          kable_styling(latex_options = c("hold_position"))
        print(tbl)
      } else {
        cat("No WQS rows found for this pollutant/outfall.\n\n")
      }

      # MF step-by-step equations
      cat("**RWC Calculation**\n\n")
      if (all(c("n_used","cv","sigma_g","p_percentile","z_p","z_target","MF") %in%
              names(sub_rp))) {
        n_val   <- sub_rp %>% dplyr::distinct(n_used)       %>% dplyr::pull() %>% .[is.finite(.)]
        cv_val  <- sub_rp %>% dplyr::distinct(cv)           %>% dplyr::pull() %>% .[is.finite(.)]
        sg_val  <- sub_rp %>% dplyr::distinct(sigma_g)      %>% dplyr::pull() %>% .[is.finite(.)]
        p_val   <- sub_rp %>% dplyr::distinct(p_percentile) %>% dplyr::pull() %>% .[is.finite(.)]
        zp_val  <- sub_rp %>% dplyr::distinct(z_p)          %>% dplyr::pull() %>% .[is.finite(.)]
        zt_val  <- sub_rp %>% dplyr::distinct(z_target)     %>% dplyr::pull() %>% .[is.finite(.)]
        mf_val  <- sub_rp %>% dplyr::distinct(MF)           %>% dplyr::pull() %>% .[is.finite(.)]
        rwc_eq  <- sub_rp %>% dplyr::distinct(RWC_rs)       %>% dplyr::pull() %>% .[is.finite(.)]
        max_val <- sub_rp %>% dplyr::distinct(max_value)    %>% dplyr::pull() %>% .[is.finite(.)]
        cl_val  <- params$confidence_level
        tgt_val <- params$target_percentile

        if (length(n_val) > 0 && length(cv_val) > 0 && length(sg_val) > 0 &&
            length(p_val) > 0 && length(zp_val) > 0 && length(zt_val) > 0 &&
            length(mf_val) > 0 && is.finite(cl_val) && is.finite(tgt_val)) {

          cat(sprintf("$$\\sigma_g = \\sqrt{\\ln(%s^2+1)} = %s$$\n",
                      format(signif(cv_val[1], 2), scientific = FALSE),
                      format(signif(sg_val[1], 2), scientific = FALSE)))
          cat(sprintf("$$p = (1 - %s)^{1/%s} = %s$$\n",
                      format(signif(cl_val, 2), scientific = FALSE),
                      format(n_val[1]),
                      format(signif(p_val[1], 2), scientific = FALSE)))
          cat(sprintf("$$z_p = %s,\\; z_{\\text{target}} = %s$$\n",
                      format(signif(zp_val[1], 2), scientific = FALSE),
                      format(signif(zt_val[1], 2), scientific = FALSE)))
          cat(sprintf(
            "$$\\mathrm{MF} = \\exp\\big((%s - %s)\\times %s\\big) = %s$$\n",
            format(signif(zt_val[1], 2), scientific = FALSE),
            format(signif(zp_val[1], 2), scientific = FALSE),
            format(signif(sg_val[1], 2), scientific = FALSE),
            format(signif(mf_val[1], 2), scientific = FALSE)))
          cat(sprintf(
            "$$\\mathrm{RWC} = \\frac{\\max(x)\\times MF}{DR} = \\frac{%s\\times%s}{%s} = %s$$\n\n",
            format(signif(max_val, 2), scientific = FALSE),
            format(signif(mf_val,  2), scientific = FALSE),
            format(signif(params$dilution_ratio, 2), scientific = FALSE),
            format(signif(rwc_eq,  2), scientific = FALSE)))
        }
      }

      # ── PLOT A: Hardness vs Limit (metals / range mode only) ────────────────
      show_range_plot <-
        (!is.null(params$hardness_mode) && params$hardness_mode == "range") &&
        any(sub_rp$CRITERION_ID %in% metals_ids &
              sub_rp$USE_CLASS_NAME_LOCATION_ETC == "class SD waters")

      if (show_range_plot) {
        crit  <- sub_rp %>%
          dplyr::filter(CRITERION_ID %in% metals_ids,
                        USE_CLASS_NAME_LOCATION_ETC == "class SD waters") %>%
          dplyr::slice(1) %>% dplyr::pull(CRITERION_ID)
        curve <- metal_limit_curve(crit)
        if (!is.null(curve) && nrow(curve) > 0) {
          ylab_txt <- sub_rp$UNIT_NAME[1]
          p_h <- ggplot2::ggplot(curve, ggplot2::aes(x = hardness, y = limit)) +
            ggplot2::geom_line(color = limit_color, linewidth = 1) +
            ggplot2::scale_y_continuous(limits = c(0, NA)) +
            ggplot2::labs(
              title = paste("Hardness vs Limit —", poll, "(Class SD)"),
              x     = "Hardness (mg/L as CaCO\u2083)",
              y     = ylab_txt
            ) +
            ggplot2::theme_minimal()
          print(p_h)
          cat("\n\n")
        }
      }

      # ── PLOT B: Observed vs WQS / RWC ───────────────────────────────────────
      # Drawing order: WQS lines (thicker, drawn first) → RWC line (on top).
      # WQS lines are drawn individually via a loop so each gets the correct
      # class colour and linetype from the shared palette.
      if (n_samples > 0 &&
          all(c("monitoring_period_end_date", "dmr_value_nmbr") %in% names(obs))) {

        ylab_txt         <- sub_rp$UNIT_NAME[1]
        present_classes  <- unique(wqs_lines$USE_CLASS_NAME_LOCATION_ETC)

        # y-axis bounds: include obs values, WQS lines, and RWC
        all_y_vals <- c(obs$dmr_value_nmbr, wqs_lines$CRITERION_VALUE,
                        if (is.finite(rwc_val)) rwc_val else numeric(0))
        y_min <- max(0, min(all_y_vals, na.rm = TRUE) * 0.95)

        p <- ggplot2::ggplot(
          obs,
          ggplot2::aes(x = monitoring_period_end_date, y = dmr_value_nmbr)
        ) +
          ggplot2::geom_line(color = "lightgrey", linetype = "dotted",
                             linewidth = 0.5)

        # Draw WQS lines first (thicker), one at a time for full palette control
        if (nrow(wqs_lines) > 0) {
          for (i in seq_len(nrow(wqs_lines))) {
            row_i <- wqs_lines[i, ]
            p <- p + ggplot2::geom_hline(
              yintercept  = row_i$CRITERION_VALUE,
              color       = wqs_colors[row_i$USE_CLASS_NAME_LOCATION_ETC],
              linetype    = wqs_linetypes[row_i$USE_CLASS_NAME_LOCATION_ETC],
              linewidth   = wqs_lwd_max,
              inherit.aes = FALSE
            )
          }
        }

        # RWC line drawn last — on top of all WQS lines
        if (is.finite(rwc_val)) {
          p <- p + ggplot2::geom_hline(
            yintercept  = rwc_val,
            color       = rwc_color,
            linetype    = "solid",
            linewidth   = rwc_lwd,
            inherit.aes = FALSE
          )
        }

        # Observation points (colour by reported unit — second colour scale)
        p <- p +
          ggplot2::geom_point(
            ggplot2::aes(color = dmr_unit_desc),
            alpha = 0.85
          ) +
          ggplot2::scale_color_manual(
            name     = "Unit",
            values   = unit_colors,
            na.value = "#999999",
            breaks   = names(unit_colors),
            drop     = FALSE
          ) +
          ggplot2::scale_y_continuous(limits = c(y_min, NA)) +
          ggplot2::labs(x = "Monitoring date", y = ylab_txt) +
          ggplot2::theme_minimal()

        # Water class legend via zero-alpha ghost layer (separate from unit scale)
        if (length(present_classes) > 0) {
          p <- p + ggnewscale::new_scale_color() +
            ggplot2::geom_hline(
              data = wqs_lines,
              ggplot2::aes(yintercept  = CRITERION_VALUE,
                           color       = USE_CLASS_NAME_LOCATION_ETC,
                           linetype    = USE_CLASS_NAME_LOCATION_ETC),
              linewidth   = wqs_lwd_max,
              inherit.aes = FALSE,
              show.legend = TRUE
            ) +
            ggplot2::scale_color_manual(
              name   = "Water Class",
              values = active_wqs_colors(present_classes)
            ) +
            ggplot2::scale_linetype_manual(
              name   = "Water Class",
              values = active_wqs_linetypes(present_classes)
            )
        }

        # Hardness annotation just above the RWC line
        if (!is.null(hardness_note) && is.finite(rwc_val)) {
          date_range <- range(obs$monitoring_period_end_date, na.rm = TRUE)
          x_pos <- date_range[1] + diff(date_range) / 2
          y_pos <- rwc_val * 1.03
          p <- p + ggplot2::annotate(
            "text", x = x_pos, y = y_pos, label = hardness_note,
            hjust = 0.5, vjust = 0, size = 3.5, color = rwc_color
          )
        }

        print(p)
        cat("\n\n")

      } else {
        cat("No dated observations to plot.\n\n")
      }

      # Detail table (WQS info per criterion)
      detail <- sub_rp %>%
        { if (!is.null(wqs_info_df))
            dplyr::left_join(., wqs_info_df, by = "CRITERION_ID") else . } %>%
        dplyr::select(
          CRITERION_ID,
          USE_CLASS_NAME_LOCATION_ETC,
          dplyr::any_of(c("CRITERIATYPEAQUAHUMHLTH", "CRITERIATYPEFRESHSALTWATER")),
          CRITERION_VALUE,
          RWC_rs,
          RP
        ) %>%
        dplyr::distinct()

      if (nrow(detail) > 0) {
        has_type_cols <- all(c("CRITERIATYPEAQUAHUMHLTH",
                               "CRITERIATYPEFRESHSALTWATER") %in% names(detail))
        col_names <- if (has_type_cols) {
          c("WQS ID", "Water Class", "Human Health Type", "Fresh/Salt Type",
            "Limit", "RWC", "RP")
        } else {
          c("WQS ID", "Water Class", "Limit", "RWC", "RP")
        }
        kable(detail,
              format    = "latex",
              booktabs  = TRUE,
              col.names = col_names) %>%
          kable_styling(latex_options = c("hold_position")) %>%
          print()
      }

    } # end outfall loop
  } # end pollutant loop
} # end dmr_df check
```


# Unavailable Pollutants {#sec-unavailable}

```{r}
#| label: unavailable
#| echo: false

# Rebuild bottom here so this chunk is self-contained and does not depend on
# the coverage chunk having run successfully (e.g. in standalone mode where
# coverage_df may be NULL).
bottom_unavail <- if (!is.null(coverage_df) && nrow(coverage_df) > 0) {
  coverage_df %>%
    dplyr::filter(`# Samples` == 0) %>%
    dplyr::arrange(Pollutant) %>%
    dplyr::select(!Form) %>%
    dplyr::distinct()
} else {
  NULL
}

if (!is.null(bottom_unavail) && nrow(bottom_unavail) > 0) {
  kable(bottom_unavail,
        format    = "latex",
        booktabs  = TRUE,
        longtable = TRUE,
        col.names = c("Pollutant", "Ref Param", "# Samples")) %>%
    kable_styling(latex_options = c("striped", "repeat_header"))
} else {
  cat("All expected pollutants had data available.")
}
```



